---
title: "GATT Analysis"
author: "Kristy Buzard"
date: 'last edited `r Sys.Date()` '
output: pdf_document
toc: yes
number_sections: yes
---

```{r setup, include=FALSE}
library(tinytex)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
```

# Next steps

## To do

1.  Create centralized documentation

    -   Include history from Unsolved problems in coding.docx (OneDrive)

2.  **Kennedy, Tokyo, Uruguay**

3.  Choose other countries

    -   Refine *Members.in.GATT.xlsx*

    -   Focus on Benelux, Canada, Chile, France, India, U.K., Dominican Republic, Haiti, Italy, Germany, Peru, Japan

        -   Matt is adding \# of pages for each schedule

4.  **Make list of accuracy checks, run them, fix typos in data**

    -   Check for tariffs going up from round to round

5.  **Figure out how to integrate "free" list**

    -   For which rounds do we have the free list typed up? Just Torquay Free List.xlsx on G: drive

6.  Condense data cleaning code

7. Go back to analysis that was thrown out ahead of CEA deadline

    - Look for gradualism in graphs

8. 10 lines in Dillon that have more than 2 years

9. Think about how variation in units affects specific summary stats

    -   Look into trade-weighting

10. TOT analysis

11. Find implementation years (maybe get answer from Doug Irwin)

12. Are current Column 2 tariffs in 1962 Smoot Hawley or the 1946 tariffs?

## Done

1.  ~~Make Github version for CEA abstract~~
2.  ~~Contact Tricia Mueller (USITC) and Roy Santana (WTO) [Bob Staiger's suggestions] [Feb 24]~~
3.  ~~Figure out how to source multiple code files~~
4.  ~~Program stats into abstract~~
5.  ~~Resolve copyright issues, then (hopefully) post the correct schedules on Github~~
6.  ~~Determine that TSUS tariffs were always at 5 digit, so we can just use the 5-digit tariff for all of the 7-digit subcategories~~
7.  ~~Read and summarize "Two Centuries of Tariffs" (USITC, in G:drive folder)~~
8.  ~~Consolidate various notes in Github / One Drive / G drive~~
9.  ~~Read and summarize "Tariff negotiations and renegotiations under the GATT and the WTO" (hard copy at SU library)~~
10.  ~~Read through Victor's notes for ideas~~
11. ~~Add Schedule A tariff data from 1946 (last available before Geneva 1947)~~
12. ~~Resolve "complicated" paragraphs, including 4 that still have no tariffs~~
13. ~~Identify lines that switch between specific and ad valorem~~
14. ~~Get working draft together ASAP~~

# Importing and cleaning the data

Importing and cleaning the data is done in "data_cleaning.rmd". It needs to be reprogrammed before being added here because it is still not as compact and readable as I want it to be. The chunk below calls that program to make the processed data available to the rest of the commands in this document.

```{r createdata,include=TRUE,message=FALSE}
source('data_cleaning.r')
```

## Sanity checks

```{r}
c <- shortnames %>%
  filter( is.na(Sp_SH) != is.na(Un_SH) & is.na(Sp_BG) != is.na(Un_BG) & is.na(Sp_Ge) != is.na(Un_Ge) &
            is.na(Sp_An) != is.na(Un_An) & is.na(Sp_To) != is.na(Un_To) & is.na(Sp_GA) != is.na(Un_GA) &
            is.na(Sp_GB) != is.na(Un_GB) & is.na(Sp_GC) != is.na(Un_GC) & is.na(Sp_DA) != is.na(Un_DA) &
            is.na(Sp_DB) != is.na(Un_DB))
```
`r nrow(c)` rows have either a specific tariff and no unit or a unit with no specific tariff for some round.

# Basic summary statistics

## Specific tariffs

We see below that the specific tariffs come down by roughly half from Smoot Hawley.

-   About half came in Geneva, the rest through Dillon. That is, Geneva did half the work and the following four rounds did the other half

But this could be deceptive since different lines use different units

-   Victor has standardized everything to be in cents (per U.S. dollar) in UnitsKey.r

```{r echo=TRUE,warning=FALSE}
source('UnitsKey.r')
```

```{r warning=FALSE,message=FALSE}
library(qwraps2)
library(kableExtra)

sp_mean <- specific[,6:15] %>% sapply(mean, na.rm=TRUE)
sp_med <- specific[,6:15] %>% sapply(median, na.rm=TRUE)
sp_count <- specific[,6:15] %>% sapply(function(x) sum(!is.na(x)))
sp_min <- specific[,6:15] %>% sapply(min, na.rm=TRUE)
sp_max <- specific[,6:15] %>% sapply(max, na.rm=TRUE)
sp_25p <- specific[,6:15] %>% sapply(function(x) quantile(x,probs=0.25,na.rm = TRUE))
sp_75p <- specific[,6:15] %>% sapply(function(x) quantile(x,probs=0.75,na.rm = TRUE))

sp_table = cbind(data.frame(sp_min),data.frame(sp_25p),data.frame(sp_mean),data.frame(sp_med),data.frame(sp_75p),data.frame(sp_max),data.frame(sp_count))
names(sp_table)<-c('Min','1st Quartile','Mean','Median','3rd Quartile','Max','N')
rownames(sp_table) <- c("Smoot Hawley", "1946","Geneva", "Annecy","Torquay","GenevaA","GenevaB","GenevaC","DillonA","DillonB")
s <- kbl(sp_table,digits=2,booktabs=T) %>%
  kable_styling(position = "center", latex_options = "hold_position")
add_header_above(s, c(" " = 1,"Summary Statistics of Specific Tariffs by Round"=7),
                 font_size = 12)
```

## Ad valorem tariffs

Strikingly, the reductions look to be of the same magnitude for Ad valorem, again with Geneva doing about half the work.

```{r echo=FALSE}
av_mean <- ad_valorem[,6:15] %>% sapply(mean, na.rm=TRUE)
av_med <- ad_valorem[,6:15] %>% sapply(median, na.rm=TRUE)
av_count <- ad_valorem[,6:15] %>% sapply(function(x) sum(!is.na(x)))
av_min <- ad_valorem[,6:15] %>% sapply(min, na.rm=TRUE)
av_max <- ad_valorem[,6:15] %>% sapply(max, na.rm=TRUE)
av_25p <- ad_valorem[,6:15] %>% sapply(function(x) quantile(x,probs=0.25,na.rm = TRUE))
av_75p <- ad_valorem[,6:15] %>% sapply(function(x) quantile(x,probs=0.75,na.rm = TRUE))

adval_table = cbind(data.frame(av_min),data.frame(av_25p),data.frame(av_mean),data.frame(av_med),data.frame(av_75p),data.frame(av_max),data.frame(av_count))
names(adval_table)<-c('Min','1st Quartile','Mean','Median','3rd Quartile','Max','N')
rownames(adval_table) <- c("Smoot Hawley","1946", "Geneva", "Annecy","Torquay","GenevaA","GenevaB","GenevaC","DillonA","DillonB")
a <- kbl(adval_table,digits=2,booktabs=T) %>%
  kable_styling(position = "center", latex_options = "hold_position")
add_header_above(a, c(" " = 1,"Summary Statistics of Ad Valorem Tariffs by Round"=7),font_size = 12)
```

-   In Dillon, `r sum(is.na(ad_valorem$Ad_Valorem_Dillon_B))` rows out of `r nrow(ad_valorem)` are missing, so there are `r nrow(ad_valorem) - sum(is.na(ad_valorem$Ad_Valorem_Dillon_B))` ad valorem tariffs. So `r round(sum(!is.na(ad_valorem$Ad_Valorem_Dillon_B))/nrow(ad_valorem)*100,2)`% of lines have *ad valorem* tariffs.

# Lines that switch between specific, ad valorem, and compound

Below are the lines that either change units or change between specific only, ad valorem only or both specific and ad valorem. Indicator variables for each round (G for Geneva, A for Annecy, etc.) show in which round the change(s) occurred. Variable "unit_ch" equals 1 if the unit changed. 

```{r}
change_BG <-
  data_set %>%
  filter(
    is.na(Specific_SH) != is.na(Specific_1946_after) | 
      is.na(Ad_Valorem_SH) != is.na(Ad_Valorem_1946_after) | Units_SH != Units_1946_after 
  ) %>%
  mutate(BG = 1,unit_ch = ifelse(Units_SH != Units_1946_after, 1, 0))
         
change_G <-
  data_set %>%
  filter(
    is.na(Specific_1946_after) != is.na(Specific_Geneva) | 
      is.na(Ad_Valorem_1946_after) != is.na(Ad_Valorem_Geneva) | Units_1946_after != Units_Geneva 
  ) %>%
  mutate(G = 1,unit_ch = ifelse(Units_1946_after != Units_Geneva, 1, 0))
         
change_A <-
  data_set %>%
  filter(
    is.na(Specific_Annecy) != is.na(Specific_Geneva) | 
      is.na(Ad_Valorem_Annecy) != is.na(Ad_Valorem_Geneva) | Units_Annecy != Units_Geneva
  ) %>%
  mutate(A = 1,unit_ch = ifelse(Units_Annecy != Units_Geneva, 1, 0))

change_T <-
  data_set %>%
  filter(
    is.na(Specific_Annecy) != is.na(Specific_Torquay) | 
      is.na(Ad_Valorem_Annecy) != is.na(Ad_Valorem_Torquay) | Units_Annecy != Units_Torquay
  ) %>%
  mutate(T = 1,unit_ch = ifelse(Units_Annecy != Units_Torquay, 1, 0))

change_GA <-
  data_set %>%
  filter(
    is.na(Specific_Geneva56_A) != is.na(Specific_Torquay) | 
      is.na(Ad_Valorem_Geneva56_A) != is.na(Ad_Valorem_Torquay) | Units_Geneva56_A != Units_Torquay
  ) %>%
  mutate(GA = 1,unit_ch = ifelse(Units_Torquay != Units_Geneva56_A, 1, 0))

change_GB <-
  data_set %>%
  filter(
    is.na(Specific_Geneva56_A) != is.na(Specific_Geneva56_B) | 
      is.na(Ad_Valorem_Geneva56_A) != is.na(Ad_Valorem_Geneva56_B) | Units_Geneva56_A != Units_Geneva56_B
  ) %>%
  mutate(GB = 1,unit_ch = ifelse(Units_Geneva56_A != Units_Geneva56_B, 1, 0))

change_GC <-
  data_set %>%
  filter(
    is.na(Specific_Geneva56_C) != is.na(Specific_Geneva56_B) | 
      is.na(Ad_Valorem_Geneva56_C) != is.na(Ad_Valorem_Geneva56_B) | Units_Geneva56_C != Units_Geneva56_B
  ) %>%
  mutate(GC = 1,unit_ch = ifelse(Units_Geneva56_B != Units_Geneva56_C, 1, 0))

change_DA <-
  data_set %>%
  filter(
    is.na(Specific_Dillon_A) != is.na(Specific_Geneva56_C) | 
      is.na(Ad_Valorem_Dillon_A) != is.na(Ad_Valorem_Geneva56_C) | Units_Dillon_A != Units_Geneva56_C
  ) %>%
  mutate(DA = 1,unit_ch = ifelse(Units_Geneva56_C != Units_Dillon_A, 1, 0))

change_DB <-
  data_set %>%
  filter(
    is.na(Specific_Dillon_A) != is.na(Specific_Dillon_B) | 
      is.na(Ad_Valorem_Dillon_A) != is.na(Ad_Valorem_Dillon_B) | Units_Dillon_A != Units_Dillon_B
  ) %>%
  mutate(DB = 1,unit_ch = ifelse(Units_Dillon_B != Units_Dillon_B, 1, 0))

mergeCols <- colnames(data_set)
mC <- mergeCols[mergeCols != "unit_ch"]
changes <- select(change_BG,-unit_ch) %>% 
  full_join(select(change_G,-unit_ch),  by = mergeCols)%>%
  full_join(select(change_A,-unit_ch),  by = mergeCols) %>%
  full_join(select(change_T,-unit_ch),  by = mergeCols) %>% 
  full_join(select(change_GA,-unit_ch),  by = mergeCols) %>%
  full_join(select(change_GB,-unit_ch),  by = mergeCols) %>% 
  full_join(select(change_GC,-unit_ch),  by = mergeCols) %>%
  full_join(select(change_DA,-unit_ch),  by = mergeCols) %>%
  full_join(select(change_DB,-unit_ch),  by = mergeCols) %>% arrange(id)
```
In all, `r nrow(changes)` lines are affected by some change in the form of the tariff.

```{r}
print.data.frame(changes[,c("Sched","Product","Paragraph","id","G","A","T","GA","GB","GC","DA","DB","Interval")],row.names = FALSE)
```


# How did liberalization vary across Schedules?

First, descriptions of each schedule:

```{r}
para_count <- data_set %>% group_by(Sched) %>% summarize(n=n()) %>% ungroup()

schedule_names<-c('Chemicals, Oil, and Paints', 'Earths, Earthenware, and Glassware', 'Metals and Manufactures of', 'Wood and Manufactures of', 'Sugar, Molasses, and Manufactures of', 'Tobacco and Manufactures of', 'Agricultural Products and Provisions', 'Spirits, Wines, and other Beverages','Cotton Manufactures','Flax, Hemp, Jute, and Manufactures of','Wool and Manufactures of', 'Silk Manufactures','Manufactures of Rayon or Other Synthetic Textile', 'Papers and Books', 'Sundries')

abbr <- c('Chem', 'Glass', 'Metals', 'Wood', 'Sugar', 'Tobacco', 'Agri', 'Spirits','Cotton','Flax','Wool', 'Silk','Rayon', 'Paper', 'Sundries')

data_schedules<-data.frame(para_count,abbr,schedule_names)
names(data_schedules)<-c('Schedule','# Lines','Abbreviation','Title')

s <- kable(data_schedules, align= 'ccl',booktabs=T) %>%
  kable_styling(full_width = F, latex_options = "hold_position")
add_header_above(s, c("Smoot Hawley Schedule Titles"=4),font_size = 12)
```

## Summary stats for specific tariffs

```{r echo=FALSE, warning=FALSE}
sum_specific_pc <- specific %>%   
  group_by(Sched) %>%  
  summarize(SH_mean = mean(Specific_SH, na.rm=TRUE),
            DB_mean = mean(Specific_Dillon_B, na.rm=TRUE),
            mean_chg = ((SH_mean - DB_mean)/SH_mean)*100,
            SH_med = median(Specific_SH, na.rm=TRUE),
            DB_med = median(Specific_Dillon_B, na.rm=TRUE),
            med_chg = ((SH_med - DB_med)/SH_med)*100,
            SH_obs = sum(!is.na(Specific_SH)),
            DB_obs = sum(!is.na(Specific_Dillon_B)),
            n = n()) %>%
  ungroup() # ungrouping variable is a good habit to prevent errors

sss <- kbl(sum_specific_pc,digits=2,booktabs=T) %>%
  kable_styling(position = "center", latex_options = "hold_position")
add_header_above(sss, c("Specific Tariffs, all items"=10),font_size = 12)

```

<!-- The table below is exactly the same as the one above EXCEPT it drops the `r sum(!is.na(data_set$Interval))` lines that are impacted by the "tax interval" issue -->

<!-- ```{r echo=FALSE, warning=FALSE} -->
<!-- sum_specific_nI_pc <- subset(specific, is.na(Interval)) %>%    -->
<!--   group_by(Sched) %>%   -->
<!--   summarize(SH_mean = mean(Specific_SH, na.rm=TRUE), -->
<!--             DB_mean = mean(Specific_Dillon_B, na.rm=TRUE), -->
<!--             mean_chg = ((SH_mean - DB_mean)/SH_mean)*100, -->
<!--             SH_med = median(Specific_SH, na.rm=TRUE), -->
<!--             DB_med = median(Specific_Dillon_B, na.rm=TRUE), -->
<!--             med_chg = ((SH_med - DB_med)/SH_med)*100, -->
<!--             SH_obs = sum(!is.na(Specific_SH)), -->
<!--             DB_obs = sum(!is.na(Specific_Dillon_B)), -->
<!--             n = n()) %>% -->
<!--   ungroup() # ungrouping variable is a good habit to prevent errors -->

<!-- kbl(sum_specific_nI_pc,digits=2,booktabs=T) %>% -->
<!--   kable_styling(position = "center", latex_options = "hold_position") -->

<!-- ``` -->

The table below is exactly the same as the one above EXCEPT it drops the `r nrow(changes)` lines that are impacted by the "switcher" issue

```{r echo=FALSE, warning=FALSE}

changes2 <- changes %>% select(id,Sched,Paragraph,Product,Interval,starts_with("Specific"))
nochanges <- rows_delete(specific, changes2, by = "id")

sum_specific_nC_pc <- nochanges %>%   
  group_by(Sched) %>%  
  summarize(SH_mean = mean(Specific_SH, na.rm=TRUE),
            DB_mean = mean(Specific_Dillon_B, na.rm=TRUE),
            mean_chg = ((SH_mean - DB_mean)/SH_mean)*100,
            SH_med = median(Specific_SH, na.rm=TRUE),
            DB_med = median(Specific_Dillon_B, na.rm=TRUE),
            med_chg = ((SH_med - DB_med)/SH_med)*100,
            SH_obs = sum(!is.na(Specific_SH)),
            DB_obs = sum(!is.na(Specific_Dillon_B)),
            n = n()) %>%
  ungroup() # ungrouping variable is a good habit to prevent errors

sss_nc <- kbl(sum_specific_nC_pc,digits=2,booktabs=T) %>%
  kable_styling(position = "center", latex_options = "hold_position")
add_header_above(sss_nc, c("Specific Tariffs, switchers removed"=10),font_size = 12)

```

Notes:

-   8 (spirits) largest, and consistent across rounds (1 ad valorem only)
-   5 (sugar) unambiguously smallest cuts, had some of the highest ad-valorem
-   Reduction in median vs. mean: split exactly half and half as to which reduction was smaller
-   Schedule 12 must be all ad valorem

### Mean of specific tariffs by schedule and round

```{r}
sum_specific_bySched <- nochanges %>%   
  group_by(Sched) %>%  
  summarize(SH = mean(Specific_SH, na.rm=TRUE),
            A = mean(Specific_1946_after, na.rm=TRUE),
            G1 = mean(Specific_Geneva, na.rm=TRUE),
            An = mean(Specific_Annecy, na.rm=TRUE),
            To = mean(Specific_Torquay, na.rm=TRUE),
            GC = mean(Specific_Geneva56_C, na.rm=TRUE),
            DB = mean(Specific_Dillon_B, na.rm=TRUE),
            chgA = ((SH - A)/SH)*100,
            chgG1 = ((A - G1)/A)*100,
            chgAn = ((G1 - An)/G1)*100,
            chgTo = ((An - To)/An)*100,
            chgGC = ((To - GC)/To)*100,
            chgDB = ((GC - DB)/GC)*100,
            ) %>%
  ungroup() # ungrouping variable is a good habit to prevent errors

sss_byround <- kbl(sum_specific_bySched,digits=2,booktabs=T) %>%
  kable_styling(position = "center", latex_options = "hold_position")
add_header_above(sss_byround, c("Mean Specific Tariffs (no changers), by round"=14),font_size = 12)

```

Removing tax interval lines

<!--```{r} -->
<!-- changes2 <- changes %>% select(id,Sched,Paragraph,Product,Interval,starts_with("Specific")) -->
<!-- nochanges <- rows_delete(specific, changes2, by = "id") -->
<!-- sum_specific_bySched_nI <- nochanges %>%  -->
<!--   group_by(Sched) %>% -->
<!--   summarize(across(starts_with("Sp"), mean, na.rm=TRUE), -->
<!--             chgGe = ((Sp_SH - Sp_Ge)/Sp_SH)*100, -->
<!--             chgAn = ((Sp_Ge - Sp_An)/Sp_Ge)*100, -->
<!--             chgTo = ((Sp_An - Sp_To)/Sp_An)*100, -->
<!--             chgGC = ((Sp_To - Sp_GC)/Sp_To)*100, -->
<!--             chgDB = ((Sp_GC - Sp_DB)/Sp_GC)*100) %>% -->
<!--   ungroup() # ungrouping variable is a good habit to prevent errors -->

<!-- sum_specific_bySched_nI %>% select(- c(Sp_GA, Sp_GB, Sp_DA))%>% -->
<!--   kbl(digits=2,booktabs=T) %>%  -->
<!--   kable_styling(position = "center") %>% -->
<!--   kable_styling(latex_options = "hold_position") -->
<!-- ``` -->

## Summary stats for ad valorem tariffs

```{r echo=FALSE, warning=FALSE}
sum_av_pc <- ad_valorem %>%   
  group_by(Sched) %>%  
  summarize(SH_mean = mean(Ad_Valorem_SH, na.rm=TRUE),
            DB_mean = mean(Ad_Valorem_Dillon_B, na.rm=TRUE),
            mean_chg = ((SH_mean - DB_mean)/SH_mean)*100,
            SH_med = median(Ad_Valorem_SH, na.rm=TRUE),
            DB_med = median(Ad_Valorem_Dillon_B, na.rm=TRUE),
            med_chg = ((SH_med - DB_med)/SH_med)*100,
            SH_obs = sum(!is.na(Ad_Valorem_SH)),
            DB_obs = sum(!is.na(Ad_Valorem_Dillon_B)),
            n = n()) %>%
  ungroup() # ungrouping variable is a good habit to prevent errors
ssav <- kbl(sum_av_pc,digits=2,booktabs=T) %>% kable_styling(position = "center") %>% 
  kable_styling(latex_options = "hold_position")
add_header_above(ssav, c("Ad valorem Tariffs, all items"=10),font_size = 12)

```

For several paragraphs, the maximum tariff for Dillon B changes when we get rid of the tax interval lines (2,9,11). Still I'm not going to print the tables with the maxes in them for now.

```{r}
changesav <- changes %>% select(id,Sched,Paragraph,Product,Interval,starts_with("Ad_Valorem"))
nochanges_av <- rows_delete(ad_valorem, changesav, by = "id")

sum_av_nC_pc <- nochanges_av %>%   
  group_by(Sched) %>% 
  summarize(SH_mean = mean(Ad_Valorem_SH, na.rm=TRUE),
            DB_mean = mean(Ad_Valorem_Dillon_B, na.rm=TRUE),
            mean_chg = ((SH_mean - DB_mean)/SH_mean)*100,
            SH_med = median(Ad_Valorem_SH, na.rm=TRUE),
            DB_med = median(Ad_Valorem_Dillon_B, na.rm=TRUE),
            med_chg = ((SH_med - DB_med)/SH_med)*100,
            SH_obs = sum(!is.na(Ad_Valorem_SH)),
            DB_obs = sum(!is.na(Ad_Valorem_Dillon_B)),
            n = n()) %>%
  ungroup() # ungrouping variable is a good habit to prevent errors
ssav_nC <- kbl(sum_av_nC_pc,digits=2,booktabs=T) %>%
  kable_styling(position = "center", latex_options = "hold_position")
add_header_above(ssav_nC, c("Ad valorem Tariffs, switchers removed"=10),font_size = 12)

```

### Mean of ad valorem tariffs by schedule and round

<!-- ```{r}
 sum_av_bySched <- ad_valorem %>%   
   group_by(Sched) %>%
   summarize(SH = mean(Ad_Valorem_SH, na.rm=TRUE),
             G1 = mean(Ad_Valorem_Geneva, na.rm=TRUE),
             An = mean(Ad_Valorem_Annecy, na.rm=TRUE),
             To = mean(Ad_Valorem_Torquay, na.rm=TRUE),
             GC = mean(Ad_Valorem_Geneva56_C, na.rm=TRUE),
             DB = mean(Ad_Valorem_Dillon_B, na.rm=TRUE),
             chgG1 = ((SH - G1)/SH)*100,
             chgAn = ((G1 - An)/G1)*100,
             chgTo = ((An - To)/An)*100,
             chgGC = ((To - GC)/To)*100,
             chgDB = ((GC - DB)/GC)*100,
             ) %>%
   ungroup() # ungrouping variable is a good habit to prevent errors

kbl(sum_av_bySched,digits=2,booktabs=T) %>%
  kable_styling(position = "center", latex_options = "hold_position")
``` -->

Removing switchers lines

```{r}
sum_av_bySched_nC <- nochanges_av %>%   
  group_by(Sched) %>%  
  summarize(SH = mean(Ad_Valorem_SH, na.rm=TRUE),
            G1 = mean(Ad_Valorem_Geneva, na.rm=TRUE),
            An = mean(Ad_Valorem_Annecy, na.rm=TRUE),
            To = mean(Ad_Valorem_Torquay, na.rm=TRUE),
            GC = mean(Ad_Valorem_Geneva56_C, na.rm=TRUE),
            DB = mean(Ad_Valorem_Dillon_B, na.rm=TRUE),
            chgG1 = ((SH - G1)/SH)*100,
            chgAn = ((G1 - An)/G1)*100,
            chgTo = ((An - To)/An)*100,
            chgGC = ((To - GC)/To)*100,
            chgDB = ((GC - DB)/GC)*100,
            ) %>%
  ungroup() # ungrouping variable is a good habit to prevent errors

ssav_byround <- kbl(sum_av_bySched_nC,digits=2,booktabs=T) %>%
  kable_styling(position = "center", latex_options = "hold_position")
add_header_above(ssav_byround, c("Mean Ad Valorem Tariffs, by round"=12),font_size = 12)

```

# What was the total reduction in negotiated tariffs under the GATT in each round?

Mean and median of specific tariffs in each round

```{r echo=FALSE, warning=FALSE, message=FALSE}

sp_mt = cbind(data.frame(sp_mean),c(0,-diff(sp_mean)/data.frame(sp_mean)[-nrow(data.frame(sp_mean)),]*100),data.frame(sp_med),c(0,-diff(sp_med)/data.frame(sp_med)[-nrow(data.frame(sp_med)),]*100))

names(sp_mt)<-c('Mean','% decrease','Median','% decrease')
rownames(sp_mt) <- c("Smoot Hawley", "1946","Geneva", "Annecy","Torquay","GenevaA","GenevaB","GenevaC","DillonA","DillonB")
x <- kbl(sp_mt,digits=2,booktabs=T) %>% 
  kable_styling(position = "center", latex_options = "hold_position")
add_header_above(x, c(" " = 1,c("Decrease in specific tariffs by round"=4)),font_size = 12)


av_mt = cbind(data.frame(av_mean),c(0,-diff(av_mean)/data.frame(av_mean)[-nrow(data.frame(av_mean)),]*100),data.frame(av_med),c(0,-diff(av_med)/data.frame(av_med)[-nrow(data.frame(av_med)),]*100))

names(av_mt)<-c('Mean','% decrease','Median','% decrease')
rownames(av_mt) <- c("Smoot Hawley", "1946","Geneva", "Annecy","Torquay","GenevaA","GenevaB","GenevaC","DillonA","DillonB")
y <- kbl(av_mt,digits=2,booktabs=T) %>%
  kable_styling(position = "center", latex_options = "hold_position")
add_header_above(y, c(" " = 1,c("Decrease in ad valorem tariffs by round"=4)),font_size = 12)

```

# Which lines were only ad valorem, only specific, or both?

## Mixed

Next we need to know about the lines that have both ad valorem and specific (or take them out from above); at least quantify them to start

```{r}
both <- data_set %>% mutate(bothSH=!is.na(data_set[,"Specific_SH"])&!is.na(data_set[,"Ad_Valorem_SH"]),bothGeneva=!is.na(data_set[,"Specific_Geneva"])&!is.na(data_set[,"Ad_Valorem_Geneva"]),bothAnnecy=!is.na(data_set[,"Specific_Annecy"])&!is.na(data_set[,"Ad_Valorem_Annecy"]),bothTorquay=!is.na(data_set[,"Specific_Torquay"])&!is.na(data_set[,"Ad_Valorem_Torquay"]),bothGeneva56A=!is.na(data_set[,"Specific_Geneva56_A"])&!is.na(data_set[,"Ad_Valorem_Geneva56_A"]),bothGeneva56B=!is.na(data_set[,"Specific_Geneva56_B"])&!is.na(data_set[,"Ad_Valorem_Geneva56_B"]),bothGeneva56C=!is.na(data_set[,"Specific_Geneva56_C"])&!is.na(data_set[,"Ad_Valorem_Geneva56_C"]),bothDillonA=!is.na(data_set[,"Specific_Dillon_A"])&!is.na(data_set[,"Ad_Valorem_Dillon_A"]),bothDillonB=!is.na(data_set[,"Specific_Dillon_B"])&!is.na(data_set[,"Ad_Valorem_Dillon_B"]))
```

How many lines have both ad valorem and specific in each round?

-   Smoot Hawley: `r sum(both$bothSH, na.rm = TRUE)`
-   Geneva 1947: `r sum(both$bothGeneva, na.rm = TRUE)`
-   Annecy: `r sum(both$bothAnnecy, na.rm = TRUE)`
-   Torquay: `r sum(both$bothTorquay, na.rm = TRUE)`
-   Geneva56A: `r sum(both$bothGeneva56A, na.rm = TRUE)`
-   Geneva56B: `r sum(both$bothGeneva56B, na.rm = TRUE)`
-   Geneva56C: `r sum(both$bothGeneva56C, na.rm = TRUE)`
-   DillonA: `r sum(both$bothDillonA, na.rm = TRUE)`
-   DillonB: `r sum(both$bothDillonB, na.rm = TRUE)`

### Victor's intuition on mixed lines

I believe many of the changes from specific tax to ad valorem or otherwise is because of the tax intervals. You could search the keywords "tax boundaries" and "tax interval(s)" in Extra column of every round to locate them.

### Proportions of specific, ad valorem, mixed

A few lines in each round had neither specific nor ad valorem. These were all fixed as of 5/15/21, but we keep this here to check in case things pop up.

```{r comment=NA}
neither <- data_set %>% mutate(neitherSH=is.na(data_set[,"Specific_SH"])&is.na(data_set[,"Ad_Valorem_SH"]),neitherAnnecy=is.na(data_set[,"Specific_Annecy"])&is.na(data_set[,"Ad_Valorem_Annecy"]),neitherDillonB=is.na(data_set[,"Specific_Dillon_B"])&is.na(data_set[,"Ad_Valorem_Dillon_B"]))

print("Smoot-Hawley")
print.data.frame(neither[which(neither$neitherSH==TRUE),c("Sched","Product","Paragraph","id")],row.names = FALSE)

# print.data.frame(neither[which(neither$neitherAnnecy==TRUE),c("Sched","Product","Paragraph","id")],row.names = FALSE)

print("Dillon B")
print.data.frame(neither[which(neither$neitherDillonB==TRUE),c("Sched","Product","Paragraph","id")],row.names = FALSE)
```

# Tariff Increases

Here we are looking round by round for lines that had an increase in either the ad valorem or specific tariff (or both). Later we will look at lines that switch from one type of tariff to the other.

```{r echo=FALSE}
increaseBG <- shortnames %>% 
  mutate(av_pc =((AV_SH - AV_BG)/AV_SH)*100,sp_pc =((Sp_SH - Sp_BG)/Sp_SH)*100) %>%
  select(Para,id,Prod,av_pc,sp_pc,AV_SH,AV_BG,Sp_SH,Sp_BG,Un_SH,Un_BG,Int) %>%
  filter(is.na(Sp_SH) == is.na(Sp_BG) & is.na(AV_SH) == is.na(AV_BG) & 
           is.na(Un_SH) == is.na(Un_BG)) %>%
  subset(sp_pc < 0 | av_pc < 0)
print("Increased tariff from Smoot Hawley to 1946 (Before Geneva)")
print.data.frame(increaseBG,row.names = FALSE,digits = 2)


increaseGe <- shortnames %>% 
  mutate(av_pc =((AV_BG - AV_Ge)/AV_BG)*100,sp_pc =((Sp_BG - Sp_Ge)/Sp_BG)*100) %>%
  select(Para,id,Prod,av_pc,sp_pc,AV_BG,AV_Ge,Sp_BG,Sp_Ge,Un_BG,Un_Ge,Int) %>%
  filter(is.na(Sp_BG) == is.na(Sp_Ge) & is.na(AV_BG) == is.na(AV_Ge) & 
           is.na(Un_BG) == is.na(Un_Ge)) %>%
  subset(sp_pc < 0 | av_pc < 0)
print("Increased tariff from 1946 to Geneva")
print.data.frame(increaseGe,row.names = FALSE,digits = 2)

increaseA <- shortnames %>% 
  mutate(av_pc =((AV_Ge - AV_An)/AV_Ge)*100,sp_pc =((Sp_Ge - Sp_An)/Sp_Ge)*100) %>%
  select(Para,id,Prod,av_pc,sp_pc,AV_Ge,AV_An,Sp_Ge,Sp_An,Un_Ge,Un_An,Int) %>%
  filter(is.na(Sp_Ge) == is.na(Sp_An) & is.na(AV_Ge) == is.na(AV_An) & 
           is.na(Un_Ge) == is.na(Un_An)) %>%
  subset(sp_pc < 0 | av_pc < 0)
print("Increased tariff from Geneva to Annecy")
print.data.frame(increaseA,row.names = FALSE,digits = 2)

increaseT <- shortnames %>%
  mutate(av_pc =((AV_An - AV_To)/AV_An)*100,sp_pc =((Sp_An - Sp_To)/Sp_An)*100) %>%
  select(Para,id,Prod,av_pc,sp_pc,AV_An,AV_To,Sp_An,Sp_To,Un_An,Un_To,Int) %>%
  filter(is.na(Sp_An) == is.na(Sp_To) & is.na(AV_An) == is.na(AV_To) & 
           is.na(Un_An) == is.na(Un_To)) %>%
  subset(sp_pc < 0 | av_pc < 0)
print("Increased tariff from Annecy to Torquay")
print.data.frame(increaseT,row.names = FALSE,digits = 2)

increaseGC <- shortnames %>%
  mutate(av_pc =((AV_To - AV_GC)/AV_To)*100,sp_pc =((Sp_To - Sp_GC)/Sp_To)*100) %>%
  select(Para,id,Prod,av_pc,sp_pc,AV_To,AV_GC,Sp_To,Sp_GC,Un_To,Un_GC,Int) %>%
  filter(is.na(Sp_To) == is.na(Sp_GC) & is.na(AV_To) == is.na(AV_GC) & 
           is.na(Un_To) == is.na(Un_GC)) %>%
  subset(sp_pc < 0 | av_pc < 0)
print("Increased tariff from Torquay to Geneva56_C")
print.data.frame(increaseGC,row.names = FALSE,digits = 2)

increaseD <- shortnames %>%
  mutate(av_pc =((AV_GC - AV_DB)/AV_GC)*100,sp_pc =((Sp_GC - Sp_DB)/Sp_GC)*100) %>%
  select(Para,id,Prod,av_pc,sp_pc,AV_GC,AV_DB,Sp_GC,Sp_DB,Un_GC,Un_DB,Int) %>%
  filter(is.na(Sp_GC) == is.na(Sp_DB) & is.na(AV_GC) == is.na(AV_DB) & 
           is.na(Un_GC) == is.na(Un_DB)) %>%
  subset(sp_pc < 0 | av_pc < 0)
print("Increased tariff from Geneva56_C to Dillon_B")
print.data.frame(increaseD,row.names = FALSE, digits = 2)
```

# No changes

## No change from Smoot Hawley to Dillon B

```{r echo=F}

same <- function(b,e){
  beg <- shortnames %>% select(ends_with(b))
  end <- shortnames %>% select(ends_with(e))
  name <- substitute(e)
  z <- as.name(paste0("same",name))
  assign(deparse(substitute(z)), shortnames %>%
           filter(((is.na(end[,3]) == T & is.na(beg[,3]) == T) | beg[,3] == end[,3])
                & ((is.na(end[,2]) == T & is.na(beg[,2]) == T) | beg[,2] == end[,2])
                & ((is.na(end[,1]) == T & is.na(beg[,1]) == T) | beg[,1] == end[,1]) )
         , envir=.GlobalEnv)
}  

same("SH","DB")

# below we get the sets of lines for each round compared to SH
same("SH","BG")
same("SH","Ge")
same("SH","An")
same("SH","To")
same("SH","GA")
same("SH","GB")
same("SH","GC")
same("SH","DA")

nr = nrow(shortnames)
num_sameSH <- c(nr,nrow(sameBG),nrow(sameGe),nrow(sameAn),
                nrow(sameTo),nrow(sameGA),nrow(sameGB),
                nrow(sameGC),nrow(sameDA),nrow(sameDB))

```

The code above produces `r nrow(sameDB)` lines that are the same in Smoot-Hawley and Dillon B (i.e. that don't change at all through these five rounds of negotiations--we assume. We still need a check for rates going up.)

## No change from Smoot Hawley to Geneva

```{r echo=F}
# we removed the "1946_before" variables once we verified that they were exactly the same as Smoot Hawley
# all the lines that are exactly the same in Smoot Hawley and 1946_before
#same <- shortnames %>%
#         filter( ((is.na(Sp_SH) == is.na(Sp_B) & is.na(Sp_SH)) | Sp_SH == Sp_B) 
#                & ((is.na(AV_SH) == is.na(AV_B) & is.na(AV_SH)) | AV_SH == AV_B)
#                & ((is.na(Un_SH) == is.na(Un_B) & is.na(Un_SH)) | Un_SH == Un_B))


# all the lines that are exactly the same in Smoot Hawley and Before Geneva
same("SH","BG")

# below we get the sets of lines for each round compared to the previou round
same("BG","Ge")
same("Ge","An")
same("An","To")
same("To","GA")
same("GA","GB")
same("GB","GC")
same("GC","DA")
same("DA","DB")

year <- c(1930,1946,1947,1949,1951,1956,1957,1958,1963,1964)
num_nego <- c(0,nr-nrow(sameBG),nr-nrow(sameGe),nr-nrow(sameAn),nr-nrow(sameTo),
            nr-nrow(sameGA),nr-nrow(sameGB),nr-nrow(sameGC),nr-nrow(sameDA),nr-nrow(sameDB))
plot1_data <- data.frame(year,num_nego,num_sameSH)

# Plot
ggplot() + 
  geom_line(data = plot1_data, aes(x=year, y= num_sameSH), color = "blue") +
  annotate("text", label = "No. of lines Still at Smoot-Hawley level", x = 1947, y = 2700, size = 5, colour = "blue") +
  geom_point(data = plot1_data, aes(x=year, y= num_nego), color = "red") +
  annotate("text", label = "No. of lines reduced since previous round", x = 1955, y = 50, size = 5, colour = "red") +
  labs(x="Year", y= NULL) +
  ggtitle("Tariff Reductions by Number of Tariff Lines") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_x_continuous(breaks = seq(1930, 1965, by = 5)) +
  scale_y_continuous(breaks = seq(0, 3000, by = 500))

```

The code above produces `r nrow(same)` lines that are the same in Smoot Hawley and Geneva.

# Summarizing the impact of tax intervals

PUT THIS BACK IN WHEN I'M AT HOME AND CAN FIGURE OUT THE BETTER WAY TO WORK WITH THE INTERVALS

```{r}

# I think I could do this natively now without having to rely on Victor's excel spreadsheet
# library(readxl)
# intervals <- read_excel("C:/Users/krist/Downloads/tax_intervals_list (1).xlsx",range = "A1:M46")

# intervals<-fill(intervals, Schedule, .direction="down")

# intervals_total <- intervals[,c("Schedule","Paragraphs_affected","Lines_affected")] %>% #group_by(Schedule) %>% summarise_all(sum) %>% ungroup()

```

# Implementation dates

Geneva 1: January 1, 1948 (Irwin 2017, p. 486)

# TOT analysis

We'll need measure of importer market power

1.  inverse foreign supply elasticities are at HS6 level, are much more recent

    -   Ross will look into the feasibility (data and code) of creating these measures for the 1930s/40s
    -   Would we want Broda, Limao, Weinstein version (requires trade flows only) or Anson Soderbery's [heterogeneous version](https://web.ics.purdue.edu/~asoderbe/Site/Elasticities.html)?
    -   Ross recalls he's seen a joint project between Anson Soderbery and Doug Irwin about the 1930s

2.  product differentiaton index (Rauch), also newer, but maybe less sensitive to changes over time

3.  market share might be credible enough, and easier to get

We'll need to think about whether it's credible to try the identification strategy Ross has used in his work
