---
title: "Quantifying GATT Trade Liberalization"
author: "Kristy Buzard, Ross Jestrab and Zeyuan (Victor) Xiong"
date: "May 20, 2021"
output:
  pdf_document:
    number_sections: true
fontsize: 12pt
linestretch: 1.25
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(english)
```

# Introduction

#```{r child = 'intro.Rmd'}
#```

# Literature

- There is a large qualitative literature on what happened through the GATT
  - Go back and list some of the books we read on, e.g., each round
  - "Tariff negotiations and renegotiations under the GATT and the WTO"
  - USITC history papers (two of them)
  - Irwin's big book
    - 1994 Irwin NBER working paper
  - Petros' chapter on tariffs


- Three strands of quantitative literature
  - Old stuff
    - Krause 1962, etc
      - Go through reference list
    - Dur, Balassa, Kreinin
  - BSY (is there a second one?)
  - Bown and Irwin (2017)

- Need to include TOT literature if going to include TOT analysis

# Data

```{r, include=FALSE}
source("C:/Users/krist/Github/Gradualism/data_cleaning.rmd", local = knitr::knit_global())
```

- Separate section first on institutional context?

- Background, general info on GATT
  - RTAA reductions between Smoot Hawley and 1946
  - Where we got these files

- How we digitized the data
  - Dealing with two changes in coding
    - Historical context on those changes
    - Addition of new lines round-by-round
  
- Other issues that arose
  - Specific vs. ad valorem vs. mixed
  - Tariff intervals
  - Free list
  - Staging (starting with Geneva A/B/C)

#```{r child = 'data.Rmd'}
#```

# Results

#```{r child = 'results.Rmd'}
#```

# Non-U.S. Contracting Parties

# Conclusion

# References

#```{r child = 'references.Rmd'}
#```

# Data Appendix

```{r child = 'data-appendix.Rmd'}
```

- include source table
  - Victor is updating to include concordance sources
  
Where will GATT participants / lengths of schedules go? A separate appendix?
