---
title: "Quantifying GATT Trade Liberalization"
author: "Kristy Buzard, Ross Jestrab and Zeyuan (Victor) Xiong"
date: "May 20, 2021"
output:
  pdf_document:
    number_sections: true
fontsize: 12pt
linestretch: 1.25
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(english)
```

# Introduction

\#`{r child = 'intro.Rmd'} #`

# Literature

-   There is a large qualitative literature on what happened through the GATT

    -   Go back and list some of the books we read on, e.g., each round

    -   "Tariff negotiations and renegotiations under the GATT and the WTO"

        -   What is Matt's status with this?

    -   USITC history papers (two of them)

    -   Irwin's big book

        -   1994 Irwin NBER working paper

    -   Petros' chapter on tariffs

-   Three strands of quantitative literature

    -   Old stuff

        -   Krause 1962, etc

            -   Go through reference list

        -   Dur, Balassa, Kreinin

    -   BSY (is there a second one?)

    -   Bown and Irwin (2017)

    -   But what about "30 years of trade policy"?

-   Need to include TOT literature if going to include TOT analysis

# Institutional Context

-   Background, general info on GATT

    -   RTAA reductions between Smoot Hawley and 1946

        -   How do the two different documents we have relate to each other

\#`{r child = 'context.Rmd'} #`

# Data

```{r, include=FALSE}
source("C:/Users/krist/Github/Gradualism/data_cleaning.rmd", local = knitr::knit_global())
```

-   Where we got these files

    -   

-   How we digitized the data

    -   Dealing with paragraph splitting, new lines

    -   Dealing with two changes in coding

        -   Historical context on those changes
        
        - The 1962 document we have represent the US tariff rate under TSUSA for the Dillon B round (I assume we will include this round data in the analysis, namely 1962->Kennedy->Tokyo)

-   Other issues that arose

    -   Specific vs. ad valorem vs. mixed
    -   Variation in units for specific tariffs, how we standardized them
    -   Tariff intervals
    -   Free list
    -   Staging (starting with Geneva A/B/C)
    -   Ad valorem equivalents (AVE) and trade weighting

```{r child = 'data.Rmd'}
```

# Results

## Basic Facts

-   Total tariff cuts from Smoot Hawley to last round we have

    -   Specific vs. ad valorem

        -   Overlap in ad valorem and specific

- How many lines were simply bound at the Smoot Hawley level
    - This is in contrast in particular to lines that are free under Smoot Hawley but do not show up in a later negotiation--they are not *bound* to be free.

-   Addition of new lines round-by-round

-   Round by round cuts (focus on end of round)

-   Magnitude and speed of liberalization across types of products

    -   Any obvious patterns to which lines have largest/smallest cuts?

## Spotlight on some interesting products

-   Round by round graphs for specific, interesting lines

    -   To do this the way I want to, need implementation dates

-   Which lines from Smoot Hawley don't see any action?

    -   Victor says: 1014, 1530e, 1544,1104a, 81, 82,318,412

-   Are there differences in which lines get staging once staging starts?

    -   Before Kennedy at least, where it was uniform...

        -   Was it *really* uniform?

-   Can we quantify which types of products get a lot of splitting of lines?

-   Metallic magnesium and metallic magnesium scrap, para 375, swiches from specific to ad valorem in Geneva56C; reduced from 50 to 45% in Dillon

    -   1102b (wools nspf) go from ad valorem in every round to specific in Dillon
    -   202.a swiches from specific in S-H to ad valorem in Geneva

-   Need to look into bicycles (para 375)--every other one gets a new specific tariff in Geneva, but was ad valorem in S-H?

-   Para 209, item 6 has tariff double in Geneva56A

    -   331 item 10 increases specific tariff from 3 to 4 in Torquay
    -   911
    -   1005.a.3 (something to do with hemp) S-H --\> Geneva unchanged; then increase

-   Para 32, "change of tax formula"? Also 202.a, 232.c,302.d,

## Other things we learned

-   How important was RTAA between 1934 and 1946?

-   What do Column 2 tariffs really represent (1962)

-   Can we test:

    -   Trade Expansion Act of 1962 provided for removing the duty on articles for which the July 1, 1962, rate was 5 percent ad valorem or less, and reducing the rates on other articles by 50 percent of the July 1, 1962, rate. (Dobson)

        -   **The major trading countries made across-the-board cuts ranging from 36 to 39 percent of previous tariff rate levels on most products**
        -   Trade Act of 1974 grants the President a 5-year negotiating authority, allowing him to reduce by as much as 60 percent any tariff rate over 5 percent in force on January 1, 1975, and also allows him to remove the duty on any article for which the existing rate is 5 percent or less.

    -   1945, the executive branch had cut in half--the maximum allowable reduction--the rates for about 40 percent of the dutiable imports

## Analysis

-   What can we say about which / why lines have ad valorem vs. specific?

    -   Is there variation over time?

-   Can we posit an explanation for the variation in both the quantity and speed of liberalization across products?

    -   Perhaps the number of countries who are negotiating?

        -   Look at correlations between the number of pages for the schedule for each country?

\#`{r child = 'results.Rmd'} #`

# Non-U.S. Contracting Parties

Need table that Matt is creating with number of pages for each schedule for each round

# Terms of Trade Analysis

At least include our plans, what data we're going to use - proof of concept using modern elasticity data would be great, even in a subset of lines

# Conclusion

Future plans

-   What role did the presence of specific tariffs, combined with inflation, have in reducing the total level of tariff protection?

    -   Need trade volume / value / price data

# References

\#`{r child = 'references.Rmd'} #`

# Data Appendix

\#`{r child = 'data-appendix.Rmd'} #`

-   overflow from data section

-   include source table

    -   Victor is updating to include concordance sources

Where will GATT participants / lengths of schedules go? A separate appendix?
