---
title: "data"
author: "Kristy Buzard"
date: 'last edited `r Sys.Date()` '
output: pdf_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


We are currently adding the Kennedy, Tokyo and Uruguay rounds for the U.S. and all rounds for several other countries. The sketch of preliminary findings reported herein is on the U.S. data on Smoot Hawley tariffs through the Dillon round.

In standardizing the tariff lines across rounds, we end up with `r nrow(data_set)` lines. The changes through time that require this standardization effort are often interesting in themselves. For instance, Paragraph 353, which covers ``electrical articles'' such as motors, tools, telegraphs, telephones and xray apparatus, is one line with a common 35 percent *ad valorem* tariff in 1930. By the end of the Dillon Round in 1964, this paragraph has been split into `r count(ad_valorem[which(ad_valorem$Paragraph == 353),])` distinct lines with tariffs ranging from `r ad_valorem %>% filter(Paragraph == 353) %>% summarize(minimum = min(Ad_Valorem_Dillon_B))`% (xray: other) to `r ad_valorem %>% filter(Paragraph == 353) %>% summarize(maximum = max(Ad_Valorem_Dillon_B))`% (wiring).

```{r, echo=FALSE}
#check these lines of code to make sure descriptions are right -- they have to be summarized so can't be hard-coded.

#data_set %>% filter(Paragraph == 353) %>% slice(which.min(Ad_Valorem_Dillon_B))
#data_set %>% filter(Paragraph == 353) %>% slice(which.max(Ad_Valorem_Dillon_B))
```

## Data sources

We have been able to locate the original documents that contain the consolidated GATT tariff schedules at the end of each of the eight round of negotiations. The GATT tariff schedule data is collected from the United Nation Treaty Collection (UNTC) website, which offers detailed tariff schedule of every country of each round of GATT negotiation from the first round of Geneva to the last Uruguay round.

We also collected data and background documents to support the process of aligning tariff lines across rounds from official U.S. government documents, some of which are publicly available and a few that we access using an institutional subscription to HathiTrust Digital Library (https://www.hathitrust.org/). Details for all data sources are in the [Data Appendix](#dataappendix).

## Digitization process

We are for now focused on the tariff schedules of United States of America. These schedules went through two systematical changes from the first round of GATT negotiation, Geneva 1947, to the last round of negotiation, Uruguay 1992. Between 1930 and 1962 the tariff classification system followed the paragraph numbers of the Tariff Act of 1930, known as "Smoot-Hawley" for the bill's sponsors.

Via the Tariff Act of 1962, the US government shifted its tariff system from Smoot-Hawley to the "Tariff Schedule of United States" (TSUS). The schedule documents became known as "Tariff Schedule of the United States Annotated" (TSUSA). In the Omnibus Trade and Competitiveness Act of 1988, the United States adopted the international Harmonized System (HS); the U.S. implementation of the system is called the Harmonized Tarif Schedule (HTS).

The classifications systems used for the various GATT round are:
- Smoot-Hawley system (five rounds): Geneva 1947, Annecy 1948, Torquay 1950, Geneva 1956 and Dillon 1960;
- TSUS system (two rounds): Kennedy 1964, Tokyo 1979;
- Harmonized system (one round): Uruguay 1992.

To date, we have digitalized and standardized the tariff data for the first five GATT rounds for the United States. We have done the same for the tariff schedule that was in effect in the United States before the start of the GATT---the so-called Smoot Hawley tariffs as well as 1946 tariff schedule that was published as the baseline for the GATT negotiations.

We have also digitized the two rounds under the TSUSA system with work on the standardization between these rounds and the earlier rounds ongoing. For more details, see [Details of the Digitization Process](#digdetails). 



## Units

Within the Smoot_Hawley tariff system, we have identified 59 types of units (based on the description showing up in the original document), we have done the normalization process to narrow down the total number to 27 units keys. The reason of normalization has two layers: first we want to keep the framework as simple as possible, second we want to make sure the values of tax rate are comparable so that statistical inference based on the specific tax rates will not be affected by the units. The normalization process contains steps including integrating units that have similar descriptive meanings, integrating units that have different basic units (cents per each, cents per 100 and cents per 1000), and finally transform all dollars based units to cents based units. The units keys system we created for Smoot-Hawley system also applies for the TSUSA system in later rounds. The detailed mappings of normalization process can be found in our data package.

## Other issues

Some other issues arose during the digitization process, including dealing with tax intervals, more specified product description and staging process for certain products facing tax reduction through multiple stages. Since these issues are mostly related to the robustness and completeness check of our data and relatively trivial, we include the detailed explanation, examples and corresponding solutions in the data appendix section.


## Concordance

We built our concordance work between the TSUSA system and Smoot-Hawley system mainly in two steps: first we construct line by line keyword match method across the two schedules we have, this gives us good matches for many lines, especially for the products that have relative short and specific description; the second step is to use the cross-reference table in the **Tariff Classification Study** file ^[See more details about this document in "data sources", data appendix]. The cross-reference table provides linkage information between the paragraphs in Smoot-Hawley and the tariff item code in TSUSA, and through this way we are able to assign TSUSA code(s) to each product in the Smoot-Hawley system schedule^[The first round of concordance work has been finished and the second round more detailed matching is still in process]. 