---
title: "data"
author: "Kristy Buzard"
date: 'last edited `r Sys.Date()` '
output: pdf_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


We are currently adding the Kennedy, Tokyo and Uruguay rounds for the U.S. and all rounds for several other countries. The sketch of preliminary findings reported herein is on the U.S. data on Smoot Hawley tariffs through the Dillon round.

In standardizing the tariff lines across rounds, we end up with `r nrow(data_set)` lines. The changes through time that require this standardization effort are often interesting in themselves. For instance, Paragraph 353, which covers ``electrical articles'' such as motors, tools, telegraphs, telephones and xray apparatus, is one line with a common 35 percent *ad valorem* tariff in 1930. By the end of the Dillon Round in 1964, this paragraph has been split into `r count(ad_valorem[which(ad_valorem$Paragraph == 353),])` distinct lines with tariffs ranging from `r ad_valorem %>% filter(Paragraph == 353) %>% summarize(minimum = min(Ad_Valorem_Dillon_B))`% (xray: other) to `r ad_valorem %>% filter(Paragraph == 353) %>% summarize(maximum = max(Ad_Valorem_Dillon_B))`% (wiring).

```{r, echo=FALSE}
#check these lines of code to make sure descriptions are right -- they have to be summarized so can't be hard-coded.

#data_set %>% filter(Paragraph == 353) %>% slice(which.min(Ad_Valorem_Dillon_B))
#data_set %>% filter(Paragraph == 353) %>% slice(which.max(Ad_Valorem_Dillon_B))
```

## Data sources

We have been able to locate the original documents that contain the consolidated GATT tariff schedules at the end of each of the eight round of negotiations. The GATT tariff schedule data is collected from the United Nation Treaty Collection (UNTC) website, which offers detailed tariff schedule of every country of each round of GATT negotiation from the first round of Geneva to the last Uruguay round.

We add to this the tariff rates from the original text of the Smoot-Hawley Act as well as those from *United States Import Duties (June 1946)*, which is a list of all the rates of duty as of that time. It had been compiled by the forerunner to the US International Trade Commission in order to give U.S. negotiators the correct baseline from which to make concessions in Geneva for the first round of the GATT.

We also collected data and background documents to support the process of aligning tariff lines across rounds from official U.S. government documents, some of which are publicly available and a few that we access using an institutional subscription to HathiTrust Digital Library (https://www.hathitrust.org/). Details for all data sources are in the [Data Appendix](#dataappendix).

## Digitization process

We are for now focused on the tariff schedules of United States of America. These schedules went through two systematical changes from the first round of GATT negotiation, Geneva 1947, to the last round of negotiation, Uruguay 1992. Between 1930 and 1962 the tariff classification system followed the paragraph numbers of the Tariff Act of 1930, known as "Smoot-Hawley" for the bill's sponsors.

Via the Tariff Act of 1962, the US government shifted its tariff system from Smoot-Hawley to the "Tariff Schedule of United States" (TSUS). The schedule documents became known as "Tariff Schedule of the United States Annotated" (TSUSA). In the Omnibus Trade and Competitiveness Act of 1988, the United States adopted the international Harmonized System (HS); the U.S. implementation of the system is called the Harmonized Tarif Schedule (HTS).

The classifications systems used for the various GATT round are:
- Smoot-Hawley system (five rounds): Geneva 1947, Annecy 1948, Torquay 1950, Geneva 1956 and Dillon 1960;
- TSUS system (two rounds): Kennedy 1964, Tokyo 1979;
- Harmonized system (one round): Uruguay 1992.

To date, we have digitalized and standardized the tariff data for the first five GATT rounds for the United States. We have done the same for the tariff schedule that was in effect in the United States before the start of the GATT---the so-called Smoot Hawley tariffs as well as 1946 tariff schedule that was published as the baseline for the GATT negotiations.

We have also digitized the two rounds under the TSUSA system with work on the standardization between these rounds and the earlier rounds ongoing. For more details, see [Details of the Digitization Process](#digdetails). 


## Concordance

We are building a concordance between the Smoot-Hawley and TSUS systems in order to create a consistent time series up through the Tokyo round. First we executed a keyword match line by line between the two. This gives us good matches for many lines, especially for the products that have relatively short and specific descriptions; The second step is to use the cross-reference table in Volume 9 of the US ITC's *Tariff Classification Study.*^[See more details about this document in "data sources", data appendix.] The cross-reference table provides linkage information between the paragraphs in Smoot-Hawley and the tariff item code in TSUSA, and we use this to assign TSUSA code(s) to each product in the Smoot-Hawley system schedule.

The second stage of the concordance building is both time and skill intensive and is still in process. When it is completed, we will move on to using the concordance between TSUS and HTS provided by the USITC^["Cross-Reference Between the Harmonized System and TSUSA Schedule" courtesy of Robert Feenstra at https://cid.econ.ucdavis.edu/usixd/wp5515d.html .] in order to integrate the Uruguay round into the dataset.


## Trade weighting and *ad valorem* equivalents (AVEs)

We are in the early stages of collecting contemporaneous data on the volume and value of trade. This requires both digitizing the data (annual data is available from 1930-1946 in the U.S. Census Bureau's "Foreign commerce and navigation of the United States" report on HathiTrust) and creating a digital concordance between the Smoot-Hawley paragraph system used to determine tariff rates and the import classification system of that era.

Once this is completed, we will be able to (a) trade-weight our summary statistics and (b) compute *ad valorem* equivalents for all specific tariffs.^[See Appedix C of Teti (2020) for a clear, concise explanation of conversion of specific tariffs into *ad valorem* equivalents.] For the time being, we analyze *ad valorem* and *specific* tariffs separately.


## Units

Any analysis of specific tariffs will be complicated by differences in units. We find that the Smoot-Hawley tariff system employed 59 types of units for specific tariffs. We have reduced the number of units to 27 through a standardization and normalization process.

The reason of normalization has two layers: first we want to keep the framework as simple as possible, second we want to make sure the values of tariff rate are comparable so that statistical inference based on the specific tariff rates will not be affected by the units. The normalization process contains steps including integrating units that have similar descriptive meanings, integrating units that have different basic units (cents per each, cents per 100 and cents per 1000), and finally transform all dollars based units to cents based units. The units keys system we created for Smoot-Hawley system also applies for the TSUSA system in later rounds. The detailed mappings of normalization process can be found in our data package.


## Other issues

Some other issues arose during the digitization process, including dealing with [tariff intervals](#int), increasing detail in product description that necessitated [splitting paragraphs](#split) and the occurrence of [multiple stages](#staging) of tariff reductions over time. These issues are mostly important for robustness and completeness of our data, so we include the detailed explanations, examples and corresponding solutions in the Data Appendix.