---
title: "Quantifying GATT Trade Liberalization"
author: "Kristy Buzard, Ross Jestrab and Zeyuan (Victor) Xiong"
date: "May 20, 2021"
output:
  bookdown::pdf_document2:
    keep_tex: true
    toc: false
    extra_dependencies: ["float"]
  pdf_document2:
    number_sections: true
fontsize: 12pt
linestretch: 1.25
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
knitr::opts_chunk$set(fig.pos = "!h", out.extra = "")
library(english)
library(bookdown)
```

# Introduction

\#`{r child = 'intro.Rmd'} #`

# Literature

```{r child = 'literature.Rmd'}
```

# Institutional Context

```{r child = 'context.Rmd'}
```

# Data

```{r, include=FALSE}
source("C:/Users/krist/Github/Gradualism/data_cleaning.r", local = knitr::knit_global())
```

```{r child = 'data.Rmd'}
```

# Results

```{r child = 'results.Rmd'}
```


# Non-U.S. Contracting Parties

Need table that Matt is creating with number of pages for each schedule for each round

-   remember that time series doesn't make sense


# Conclusion

Future plans

-   What role did the presence of specific tariffs, combined with inflation, have in reducing the total level of tariff protection?

    -   Need trade volume / value / price data
    
    -   include our plans for TOT analysis, what data we're going to use - proof of concept using modern elasticity data would be great, even in a subset of lines


# References

\#`{r child = 'references.Rmd'} #`

# Data Appendix{#dataappendix}

```{r child = 'data-appendix.Rmd'}
```
