---
title: "Quantifying U.S. GATT Trade Liberalization"
author:
  - Kristy Buzard^[Syracuse University, kbuzard@syr.edu, http://kristybuzard.com]
  - Ross Jestrab^[Syracuse University, rcjestra@syr.edu]
  - Zeyuan (Victor) Xiong^[Syracuse University, zxiong07@syr.edu ]
date: "May 20, 2021"
output:
  bookdown::pdf_document2:
    keep_tex: true
    toc: false
    extra_dependencies: ["float"]
  pdf_document2:
    number_sections: true
fontsize: 12pt
linestretch: 1.25
urlcolor: blue
abstract: Large-scale tariff reductions within the framework of the General Agreement on Tariffs and Trade (GATT) have come as a result of a series of eight rounds of negotiations that began in 1947. Bown and Irwin (2017) have documented the available data---all applied tariffs---and found it to be frustratingly sparse. Somewhat surprisingly, data on the tariff commitments made by the GATT signees has not been widely available, preventing empirical studies on the workings of the GATT and the dynamics of this important episode of trade liberalization. We have been able to locate the original documents that contain the consolidated tariff schedules for each GATT round. To date, we have digitalized and standardized the tariff data for the first five rounds for the United States as well as the Smoot Hawley tariffs that were in effect in the United States before the start of the GATT. Work is underway to add the Kennedy, Tokyo and Uruguay rounds for the U.S. and all rounds for several other countries. We summarize this rich data and study the patterns across time and industries and look for explanations for these patterns.
subtitle: Preliminary; please do not circulate.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
knitr::opts_chunk$set(fig.pos = "!h", out.extra = "")
library(english)
library(bookdown)
```

[Click for latest version](https://github.com/kbuzard-SU/gradualism/blob/master/data-paper/data-paper.pdf)

# Introduction

```{r child = 'intro.Rmd'}
```

## Literature

```{r child = 'literature.Rmd'}
```

# Institutional Context

```{r child = 'context.Rmd'}
```

# Data

```{r, include=FALSE}
source("C:/Users/krist/Github/Gradualism/data_cleaning.r", local = knitr::knit_global())
```

```{r child = 'data.Rmd'}
```

# Results

```{r child = 'results.Rmd'}
```


# Non-U.S. Contracting Parties

Need table that Matt is creating with number of pages for each schedule for each round

-   remember that time series doesn't make sense


# Conclusion

Future plans

-   What role did the presence of specific tariffs, combined with inflation, have in reducing the total level of tariff protection?

    -   Need trade volume / value / price data
    
    -   include our plans for TOT analysis, what data we're going to use - proof of concept using modern elasticity data would be great, even in a subset of lines


# References

```{r child = 'references.Rmd'}
```

# Data Appendix{#dataappendix}

```{r child = 'data-appendix.Rmd'}
```
