---
title: "Quantifying GATT Trade Liberalization"
author: "Kristy Buzard, Ross Jestrab and Zeyuan (Victor) Xiong"
date: "May 20, 2021"
output:
  pdf_document:
    number_sections: true
fontsize: 12pt
linestretch: 1.25
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(english)
```

# Introduction

\#`{r child = 'intro.Rmd'} #`

# Literature

```{r child = 'literature.Rmd'}
```

# Institutional Context

-   Background, general info on GATT

    -   RTAA reductions between Smoot Hawley and 1946

        -   How do the two different documents we have relate to each other

\#`{r child = 'context.Rmd'} #`

# Data

```{r, include=FALSE}
source("C:/Users/krist/Github/Gradualism/data_cleaning.rmd", local = knitr::knit_global())
```

```{r child = 'data.Rmd'}
```

# Results

## Basic Facts

-   Total tariff cuts from Smoot Hawley to last round we have

    -   Specific vs. ad valorem

        -   Overlap in ad valorem and specific

        -   Specific vs. ad valorem vs. compound (both specific and ad valorem)

          -   Teti (2020) reports that 8% of U.S. tariffs were specific in 1988-2017
          -   Teti (2020) also reports: "Mixed tariffs are expressed as either a specific or an ad valorem rate, depending on which generates the most (or sometimes the least) revenue. Then there are technical tariffs that depend on certain product characteristics for example duties might be 8% for butter with fat content between 9-40%. Tariff rate quotas are made up of a low tariff rate on the initial imports (the within-quota quantity) and a very high tariff rate on imports entering above the initial amount (outside-quota quantity).

-   How many lines were simply bound at the Smoot Hawley level

    -   This is in contrast in particular to lines that are free under Smoot Hawley but do not show up in a later negotiation--they are not *bound* to be free.

-   Addition of new lines round-by-round

-   Round by round cuts (focus on end of round)

-   Magnitude and speed of liberalization across types of products

    -   Any obvious patterns to which lines have largest/smallest cuts?

## Spotlight on some interesting products

-   Round by round graphs for specific, interesting lines

    -   To do this the way I want to, need implementation dates

-   Which lines from Smoot Hawley don't see any action?

    -   Victor says: 1014, 1530e, 1544,1104a, 81, 82,318,412

-   Are there differences in which lines get staging once staging starts?

    -   Before Kennedy at least, where it was uniform...

        -   Was it *really* uniform?

-   Can we quantify which types of products get a lot of splitting of lines?

-   Metallic magnesium and metallic magnesium scrap, para 375, swiches from specific to ad valorem in Geneva56C; reduced from 50 to 45% in Dillon

    -   1102b (wools nspf) go from ad valorem in every round to specific in Dillon
    -   202.a swiches from specific in S-H to ad valorem in Geneva

-   Need to look into bicycles (para 375)--every other one gets a new specific tariff in Geneva, but was ad valorem in S-H?

-   Para 209, item 6 has tariff double in Geneva56A

    -   331 item 10 increases specific tariff from 3 to 4 in Torquay
    -   911
    -   1005.a.3 (something to do with hemp) S-H --\> Geneva unchanged; then increase

-   Para 32, "change of tax formula"? Also 202.a, 232.c,302.d,

## Other things we learned

-   How important was RTAA between 1934 and 1946?

-   What do Column 2 tariffs really represent (1962)

-   Can we test:

    -   Trade Expansion Act of 1962 provided for removing the duty on articles for which the July 1, 1962, rate was 5 percent ad valorem or less, and reducing the rates on other articles by 50 percent of the July 1, 1962, rate. (Dobson)

        -   **The major trading countries made across-the-board cuts ranging from 36 to 39 percent of previous tariff rate levels on most products**
        -   Trade Act of 1974 grants the President a 5-year negotiating authority, allowing him to reduce by as much as 60 percent any tariff rate over 5 percent in force on January 1, 1975, and also allows him to remove the duty on any article for which the existing rate is 5 percent or less.

    -   1945, the executive branch had cut in half--the maximum allowable reduction--the rates for about 40 percent of the dutiable imports

## Analysis

-   What can we say about which / why lines have ad valorem vs. specific?

    -   Is there variation over time?

-   Can we posit an explanation for the variation in both the quantity and speed of liberalization across products?

    -   Perhaps the number of countries who are negotiating?

        -   Look at correlations between the number of pages for the schedule for each country?

\#`{r child = 'results.Rmd'} #`

# Non-U.S. Contracting Parties

Need table that Matt is creating with number of pages for each schedule for each round

-   remember that time series doesn't make sense

# Terms of Trade Analysis

At least include our plans, what data we're going to use - proof of concept using modern elasticity data would be great, even in a subset of lines

# Conclusion

Future plans

-   What role did the presence of specific tariffs, combined with inflation, have in reducing the total level of tariff protection?

    -   Need trade volume / value / price data

# References

\#`{r child = 'references.Rmd'} #`

# Data Appendix{#dataappendix}

```{r child = 'data-appendix.Rmd'}
```
