---
title: "Quantifying GATT Trade Liberalization"
author: "Kristy Buzard, Ross Jestrab and Zeyuan (Victor) Xiong"
date: "May 20, 2021"
output:
  pdf_document:
    number_sections: true
fontsize: 12pt
linestretch: 1.25
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(english)
```

# Introduction

#```{r child = 'intro.Rmd'}
#```

# Literature

- There is a large qualitative literature on what happened through the GATT
  - Go back and list some of the books we read on, e.g., each round
  - "Tariff negotiations and renegotiations under the GATT and the WTO"
  - USITC history papers (two of them)
  - Irwin's big book
    - 1994 Irwin NBER working paper
  - Petros' chapter on tariffs


- Three strands of quantitative literature
  - Old stuff
    - Krause 1962, etc
      - Go through reference list
    - Dur, Balassa, Kreinin
  - BSY (is there a second one?)
  - Bown and Irwin (2017)
  - But what about "30 years of trade policy"?

- Need to include TOT literature if going to include TOT analysis

# Institutional Context

- Background, general info on GATT
  - RTAA reductions between Smoot Hawley and 1946
    - How do the two different documents we have relate to each other

#```{r child = 'context.Rmd'}
#```

# Data

```{r, include=FALSE}
source("C:/Users/krist/Github/Gradualism/data_cleaning.rmd", local = knitr::knit_global())
```

- Where we got these files

- How we digitized the data
  - Dealing with two changes in coding
    - Historical context on those changes
    - Addition of new lines round-by-round
  
- Other issues that arose
  - Specific vs. ad valorem vs. mixed
  - Variation in units for specific tariffs, how we standardized them
  - Tariff intervals
  - Free list
  - Staging (starting with Geneva A/B/C)
  - Ad valorem equivalents (AVE) and trade weighting

#```{r child = 'data.Rmd'}
#```

# Results

## Kristy's to-do for April 22-23

- Read through the old notes that Victor sent
- Read through my notes from the background reading for questions we want to address analytically
- Look at Plan.docx questions

## Basic Facts

- Total tariff cuts from Smoot Hawley to last round we have
  - Specific vs. ad valorem
    - Overlap in ad valorem and specific

- Round by round cuts (focus on end of round)

- Magnitude and speed of liberalization across types of products

## Spotlight on some interesting products

- Round by round graphs for specific, interesting lines
  - To do this the way I want to, need implementation dates

- Which lines from Smoot Hawley don't see any action?

- Are there differences in which lines get staging once that happens?
  - Before Kennedy at least, where it was uniform...
    - Was it *really* uniform?
    
- Can we quantify which types of products get a lot of splitting of lines?

## Other things we learned

- How important was RTAA between 1934 and 1946?

- What do Column 2 tariffs really represent (1962)

- Can we test:
  - Trade Expansion Act of 1962 provided for removing the duty on articles for which the July 1, 1962, rate was 5 percent ad valorem or less, and reducing the rates on other articles by 50 percent of the July 1, 1962, rate. (Dobson)
    - **The major trading countries made across-the-board cuts ranging from 36 to 39 percent of previous tariff rate levels on most products**
	- Trade Act of 1974 grants the President a 5-year negotiating authority, allowing him to reduce by as much as 60 percent any tariff rate over 5 percent in force on January 1, 1975, and also allows him to remove the duty on any article for which the existing rate is 5 percent or less.
- 1945, the executive branch had cut in half--the maximum allowable reduction--the rates for about 40 percent of the dutiable imports


## Analysis

- What can we say about which / why lines have ad valorem vs. specific?
  - Is there variation over time?

- Can we posit an explanation for the variation in both the quantity and speed of liberalization across products?

#```{r child = 'results.Rmd'}
#```

# Non-U.S. Contracting Parties

Need table that Matt is creating with number of pages for each schedule for each round

# Terms of Trade Analysis

At least include our plans, what data we're going to use
- proof of concept using modern elasticity data would be great, even in a subset of lines

# Conclusion

# References

#```{r child = 'references.Rmd'}
#```

# Data Appendix

```{r child = 'data-appendix.Rmd'}
```

- include source table
  - Victor is updating to include concordance sources
  
Where will GATT participants / lengths of schedules go? A separate appendix?
