---
title: "Quantifying U.S. GATT Trade Liberalization"
author:
  - Kristy Buzard^[Syracuse University, kbuzard@syr.edu, http://kristybuzard.com]
  - Ross Jestrab^[Syracuse University, rcjestra@syr.edu]
  - Zeyuan (Victor) Xiong^[Syracuse University, zxiong07@syr.edu ]
date: "May 20, 2021"
output:
  bookdown::pdf_document2:
    keep_tex: true
    toc: false
    extra_dependencies: ["float"]
  pdf_document2:
    number_sections: true
fontsize: 12pt
linestretch: 1.25
urlcolor: blue
abstract: Large-scale tariff reductions within the framework of the General Agreement on Tariffs and Trade (GATT) have come as a result of a series of eight rounds of negotiations that began in 1947. Bown and Irwin (2017) have documented the available data---all applied tariffs---and found it to be frustratingly sparse. Somewhat surprisingly, data on the tariff commitments made by the GATT signees has not been widely available, preventing empirical studies on the workings of the GATT and the dynamics of this important episode of trade liberalization. We have been able to locate the original documents that contain the consolidated tariff schedules for each GATT round. To date, we have digitalized and standardized the tariff data for the first five rounds for the United States as well as the Smoot Hawley tariffs that were in effect in the United States before the start of the GATT. Work is underway to add the Kennedy, Tokyo and Uruguay rounds for the U.S. and all rounds for several other countries. We summarize this rich data and study the patterns across time and industries.
subtitle: Preliminary; please do not circulate.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
knitr::opts_chunk$set(fig.pos = "!h", out.extra = "")
library(english)
library(bookdown)
```

[Click for latest version](https://github.com/kbuzard-SU/gradualism/blob/master/data-paper/data-paper.pdf)

# Introduction

```{r child = 'intro.Rmd'}
```

## Literature

```{r child = 'literature.Rmd'}
```

# Institutional Context {#context}

```{r child = 'context.Rmd'}
```

# Data {#data}

```{r, include=FALSE}
source("C:/Users/krist/Github/Gradualism/data_cleaning.r", local = knitr::knit_global())
```

```{r child = 'data.Rmd'}
```

# Results {#results}

```{r child = 'results.Rmd'}
```


<!-- # Non-U.S. Contracting Parties -->

<!-- Need table that Matt is creating with number of pages for each schedule for each round -->

<!-- -   remember that time series doesn't make sense -->


# Conclusion

This is the first draft of this paper and leaves room for much more analysis to be added. Adding data on *ad valorem* equivalents of the specific tariffs will allow a more unified approach. Adding price and volume data will also allow us to study to extent to which these negotiated reductions---versus inflation---impacted overall trade barriers.

The addition of these data along with breaking down the tariff changes from 1934 to 1946 to attribute to the country with whom the tariff was negotiated  will also allow us to add an analysis of the terms-of-trade theory of trade agreements.

# References

```{r child = 'references.Rmd'}
```

\newpage
# Data Appendix{#dataappendix}

```{r child = 'data-appendix.Rmd'}
```
