---
title: "Quantifying U.S. GATT Trade Liberalization"
author: |
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  beamer_presentation:
    theme: "Ilmenau"
    slide_level: 3
header-includes:
  - \setbeamertemplate{navigation symbols}{}
  - \AtBeginDocument{\title[Quantifying U.S. GATT Trade Liberalization  \hspace{2.95in}\insertframenumber/\inserttotalframenumber]{Quantifying U.S. GATT Trade Liberalization}}
  - \AtBeginDocument{\author[Buzard (kbuzard@syr.edu), Jestrab and Xiong]{Kristy Buzard, Syracuse University, kbuzard@syr.edu\\ Ross Jestrab, Syracuse University\\ Zeyuan (Victor) Xiong, Syracuse University}}
  - \usecolortheme{beaver}
  - \AtBeginSection{}
  - \AtBeginSubsection{}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Overview

- Bullet 1
- Bullet 2
- Bullet 3

# Background

- Bullet 1
- Bullet 2
- Bullet 3

# Data

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```


# Results

## Basic Facts

### Item level

- Graph

### Liberalization from 1930 to 1964

```{r, cars2, fig.cap="A scatterplot.", echo=FALSE}
plot(cars)
```

### Round-by-Round liberalization

### Industry-by-industry liberalization

- Schedule titles
- Reductions by schedule

## Notable Findings

### Importance of pre-GATT negotiations

### Some lines see tariff increases

### Very few changes between *ad valorem* and specific

# Next steps

## What's next?

- Concordances
  - Smoot-Hawley to TSUS
  - TSUS to HS
  - Smoot-Hawley to 1930's import classification system
- Import volume and value data
  - *Ad valorem* equivalents
  - Trade weighting
  - Terms-of-trade analysis