---
title: "Quantifying U.S. GATT Trade Liberalization"
author: |
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  beamer_presentation:
    theme: "Ilmenau"
    slide_level: 3
header-includes:
  - \AtBeginDocument{\setbeamertemplate{itemize item}[triangle]}
  - \AtBeginDocument{\setbeamertemplate{itemize subitem}[triangle]}
  - \AtBeginDocument{\setbeamercolor{item}{fg=black}}
  - \setbeamertemplate{navigation symbols}{}
  - \AtBeginDocument{\title[Quantifying U.S. GATT Trade Liberalization  \hspace{2.60in}\insertframenumber/\inserttotalframenumber]{Quantifying U.S. GATT Trade Liberalization}}
  - \AtBeginDocument{\author[Buzard (kbuzard@syr.edu), Jestrab and Xiong]{Kristy Buzard, Syracuse University, kbuzard@syr.edu\\ Ross Jestrab, Syracuse University\\ Zeyuan (Victor) Xiong, Syracuse University}}
  - \usecolortheme{beaver}
  - \AtBeginSection{}
  - \AtBeginSubsection{}
  - \usepackage{booktabs}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE}
source("C:/Users/krist/Github/Gradualism/data_cleaning.r", local = knitr::knit_global())
```

# Overview

Large-scale tariff reductions under GATT 1947-1994

  - 8 rounds of negotiations 

  - Negotiated tariff bindings (vs. applied tariffs)

Largely theoretical literature because no systematic data on either applied or negotiated tariffs (Bown and Irwin 2017)

  - Therefore LOTS of open research questions

    - One exception: Bagwell, Staiger and Yurukoglu (2020)

\vskip.1in
Today: describe the tariff schedule data, its sources, our methods for digitizing it,  preliminary findings for first five rounds

  
# Background

### Pre-GATT

We take advantage of a large qualitative literature on historical trade policy (e.g., Irwin 2017)

- contextualize and interpret the data, find new data sources

\vskip.1in
For the United States: 

- 1930: Smoot-Hawley Act: massive upward revision of tariffs

- 1934: Backlash from Smoot-Hawley $\Rightarrow$ Reciprocal Trade Agreements Act (RTAA)

  - President could reduce tariffs up to 50% without Congress

  - U.S. made about two dozen trade agreements by 1946

- 1945: Truman granted authority for 50% reductions from 1945 rates



### GATT era

1947, **Geneva**: bilateral negotiations $\Rightarrow$ MFN

1949, **Annecy**: only negotiate with newcomers, peril points (pp)

1951, **Torquay**: no peril points

1956, **Geneva 1956**: pp, 15% reduction allowed, 3 phases

1962, **Dillon**: pp, 20% reduction allowed, 2 phases

  - **Kennedy** begins, 50% reductions authorized

1963: New nomenclature (TSUS) takes effect


### GATT era

1967, **Kennedy**: first reductions by formula, 5 phases

\vskip.075in
1973: **Tokyo** begins, no negotiating authority for Nixon

  - 1974: granted 60% reductions, Fast Track (now TPA)
  - 1979: **Tokyo** concluded, reductions again by formula

\vskip.075in
1986: **Uruguay** starts, no negotiating authority for Reagan

  - 1988: HS replaces TSUS
  - 1993: **Uruguay** concluded


# Data

### Data sources

- GATT schedules available in PDF format in the United National Treaty Collection (UNTC)

- Tariff Act of 1930 (Smoot Hawley) from Congressional records

- *United States Import Duties, 1964*, USITC via HathiTrust Digital Library

- *Tariff Classification Study, Vol. 9*, United States Tariff Commission

- *Tariff Schedules of the United States* (TSUS), United States Tariff Commission


### So far...

To date, we've digitized and standardized the tariff schedules for the United States from 1930 to 1964, including 1946

- We expand the Smoot-Hawley nomenclature with new 'items' as sub-divisions are added through time

- About 15% of items are "compound" (Teti 2020), i.e. have both *ad valorem* and specific components 

<!-- 475 to 505 -->

- Roughly 10% of items are "mixed", i.e. have either an *ad valorem* or specific rate, depending on which is higher

- About 2% of the items are "technical", e.g. defined on proportion of content that meets some criteria

<!-- We do not analyze these categories separately because, insofar as possible, we have integrated them with the basic *ad valorem* and specific types -->

All are included in the statistics below for specific and *ad valorem*


### Still to be done

- Concordances
  - Smoot-Hawley to TSUS
  - TSUS to HS
  - Smoot-Hawley to 1930's import classification system
- Add final three GATT rounds
  - Integrate "Free Lists" (Schedule 15, ~200 items by Torquay)
- Import volume and value data
  - *Ad valorem* equivalents
  - Trade weighting
  - Terms-of-trade analysis
- Other countries' schedules


### That means...

Statistics throughout the paper for specific tariffs are *not* trade weighted; we are in the process of acquiring the trade data required to both trade weight summary statistics and compute *ad valorem* equivalents.

We analyze specific and *ad valorem* separately

We have to deal with units for specific tariffs

- Put everything in cents

- Translate tariffs to align with a few units (¢/ounce, ¢/sq. foot, ¢/each, ¢/board foot, ¢/cubic foot)


# Basic Facts

### Item level
```{r echo=F}
same <- function(b,e){
  beg <- shortnames %>% select(ends_with(b))
  end <- shortnames %>% select(ends_with(e))
  name <- substitute(e)
  z <- as.name(paste0("same",name))
  assign(deparse(substitute(z)), shortnames %>%
           filter(((is.na(end[,3]) == T & is.na(beg[,3]) == T) | beg[,3] == end[,3])
                & ((is.na(end[,2]) == T & is.na(beg[,2]) == T) | beg[,2] == end[,2])
                & ((is.na(end[,1]) == T & is.na(beg[,1]) == T) | beg[,1] == end[,1]) )
         , envir=.GlobalEnv)
}  

# below we get the sets of lines for each round compared to SH
same("SH","DB")
same("SH","BG")
same("SH","Ge")
same("SH","An")
same("SH","To")
same("SH","GA")
same("SH","GB")
same("SH","GC")
same("SH","DA")
same("SH","DB")

nr = nrow(shortnames)
num_sameSH <- c(nr,nrow(sameBG),nrow(sameGe),nrow(sameAn),
                nrow(sameTo),nrow(sameGA),nrow(sameGB),
                nrow(sameGC),nrow(sameDA),nrow(sameDB))

same2 <- function(b,e){
  beg <- shortnames %>% select(ends_with(b))
  end <- shortnames %>% select(ends_with(e))
  name1 <- substitute(b)
  name2 <- substitute(e)
  z <- as.name(paste0("same",name1,"_",name2))
  assign(deparse(substitute(z)), shortnames %>%
           filter(((is.na(end[,3]) == T & is.na(beg[,3]) == T) | beg[,3] == end[,3])
                & ((is.na(end[,2]) == T & is.na(beg[,2]) == T) | beg[,2] == end[,2])
                & ((is.na(end[,1]) == T & is.na(beg[,1]) == T) | beg[,1] == end[,1]) )
         , envir=.GlobalEnv)
}  

# all the lines that are exactly the same in Smoot Hawley and Before Geneva
same2("SH","BG")

# below we get the sets of lines for each round compared to the previou round
same2("BG","Ge")
same2("Ge","An")
same2("An","To")
same2("To","GA")
same2("GA","GB")
same2("GB","GC")
same2("GC","DA")
same2("DA","DB")

year <- c(1930,1946,1947,1949,1951,1956,1957,1958,1963,1964)
num_nego <- c(0,nr-nrow(sameSH_BG),nr-nrow(sameBG_Ge),nr-nrow(sameGe_An),nr-nrow(sameAn_To),
            nr-nrow(sameTo_GA),nr-nrow(sameGA_GB),nr-nrow(sameGB_GC),
            nr-nrow(sameGC_DA),nr-nrow(sameDA_DB))
plot1_data <- data.frame(year,num_nego,num_sameSH)
```


```{r, echo=FALSE, out.width="75%", fig.cap=""}
knitr::include_graphics("data-paper_files/figure-latex/ibi-1.pdf")
```


### Liberalization from 1930 to 1964
```{r echo=FALSE, warning=FALSE}
av_med <- data.frame(ad_valorem[,6:15] %>% sapply(median, na.rm=TRUE))
av_mean <- data.frame(ad_valorem[,6:15] %>% sapply(mean, na.rm=TRUE))
av_count <- data.frame(ad_valorem[,6:15] %>% sapply(function(x) sum(!is.na(x))))
av_min <- data.frame(ad_valorem[,6:15] %>% sapply(min, na.rm=TRUE))
av_max <- data.frame(ad_valorem[,6:15] %>% sapply(max, na.rm=TRUE))
av_25p <- data.frame(ad_valorem[,6:15] %>% sapply(function(x) quantile(x,probs=0.25,na.rm = TRUE)))
av_75p <- data.frame(ad_valorem[,6:15] %>% sapply(function(x) quantile(x,probs=0.75,na.rm = TRUE)))

sp_mean <- data.frame(specific[,6:15] %>% sapply(mean, na.rm=TRUE))
sp_med <- data.frame(specific[,6:15] %>% sapply(median, na.rm=TRUE))
sp_count <- data.frame(specific[,6:15] %>% sapply(function(x) sum(!is.na(x))))
sp_min <- data.frame(specific[,6:15] %>% sapply(min, na.rm=TRUE))
sp_max <- data.frame(specific[,6:15] %>% sapply(max, na.rm=TRUE))
sp_25p <- data.frame(specific[,6:15] %>% sapply(function(x) quantile(x,probs=0.25,na.rm = TRUE)))
sp_75p <- data.frame(specific[,6:15] %>% sapply(function(x) quantile(x,probs=0.75,na.rm = TRUE)))
```

From the Smoot-Hawley tariffs (1930) to the Dillon Round (1964) both *ad valorem* and specific tariffs were cut roughly in half

- mean *ad valorem* tariff binding decreases from `r round(av_mean["Ad_Valorem_SH",],1)`% to `r round(av_mean["Ad_Valorem_Dillon_B",],1)`%

  - medians drop from `r round(av_med["Ad_Valorem_SH",],0)`% to `r round(av_med["Ad_Valorem_Dillon_B",],0)`%

    <!-- Roughly two-thirds of the items in our dataset have an *ad valorem* tariff. -->

- mean specific tariff binding decreases from `r round(sp_mean["Specific_SH",],0)`¢ to `r round(sp_mean["Specific_Dillon_B",],0)`¢

  - medians are much smaller, dropping from `r round(sp_med["Specific_SH",],2)`¢ to `r round(sp_med["Specific_Dillon_B",],2)`¢

  <!-- Specific tariffs are present for roughly half of the items in our data.  -->





## Round-by-Round liberalization

### Table 1

```{r table_1}
sp_mean <- specific[,6:15] %>% sapply(mean, na.rm=TRUE)
sp_med <- specific[,6:15] %>% sapply(median, na.rm=TRUE)
av_mean <- ad_valorem[,6:15] %>% sapply(mean, na.rm=TRUE)
av_med <- ad_valorem[,6:15] %>% sapply(median, na.rm=TRUE)

sp_mt = cbind(data.frame(sp_mean),c(0,-diff(sp_mean)/data.frame(sp_mean)[-nrow(data.frame(sp_mean)),]*100),data.frame(sp_med),c(0,-diff(sp_med)/data.frame(sp_med)[-nrow(data.frame(sp_med)),]*100))

names(sp_mt)<-c('Mean','% decrease','Median','% decrease')
rownames(sp_mt) <- c("Smoot Hawley", "1946","Geneva", "Annecy","Torquay","GenevaA","GenevaB","GenevaC","DillonA","DillonB")

kbl(sp_mt,digits=2,booktabs=T, caption = "Decrease in Specific Tariffs by Round") %>%
  kable_styling(position = "center", latex_options = "hold_position",font_size=7)
```


### Table 2
```{r table_2}
av_mt = cbind(data.frame(av_mean),c(0,-diff(av_mean)/data.frame(av_mean)[-nrow(data.frame(av_mean)),]*100),data.frame(av_med),c(0,-diff(av_med)/data.frame(av_med)[-nrow(data.frame(av_med)),]*100))

names(av_mt)<-c('Mean','% decrease','Median','% decrease')
rownames(av_mt) <- c("Smoot Hawley", "1946","Geneva", "Annecy","Torquay","GenevaA","GenevaB","GenevaC","DillonA","DillonB")

kbl(av_mt,digits=2,booktabs=T, caption = "Decrease in Ad Valorem Tariffs by Round") %>%
  kable_styling(position = "center", latex_options = "hold_position",font_size=7)
```


### Table 3
```{r table_3}
sp_table = cbind(data.frame(sp_min),data.frame(sp_25p),data.frame(sp_mean),data.frame(sp_med),data.frame(sp_75p),data.frame(sp_max),data.frame(sp_count))
names(sp_table)<-c('Min','1st Quartile','Mean','Median','3rd Quartile','Max','N')
rownames(sp_table) <- c("Smoot Hawley", "1946","Geneva", "Annecy","Torquay","GenevaA","GenevaB","GenevaC","DillonA","DillonB")

kbl(sp_table,digits=2,booktabs=T,caption = "Summary Statistics for Specific Tariffs by Round") %>%
  kable_styling(position = "center", latex_options = "hold_position", font_size=6)
```



### Table 4
```{r table_4}
adval_table = cbind(data.frame(av_min),data.frame(av_25p),data.frame(av_mean),data.frame(av_med),data.frame(av_75p),data.frame(av_max),data.frame(av_count))
names(adval_table)<-c('Min','1st Quartile','Mean','Median','3rd Quartile','Max','N')
rownames(adval_table) <- c("Smoot Hawley","1946", "Geneva", "Annecy","Torquay","GenevaA","GenevaB","GenevaC","DillonA","DillonB")
kbl(adval_table,digits=2,booktabs=T,caption = "Summary Statistics of Ad Valorem Tariffs by Round") %>%
  kable_styling(position = "center", latex_options = "hold_position",font_size=6)
```


## Industry-by-industry liberalization

### Schedule Titles
```{r schedule_titles, message=F}
para_count <- data_set %>% group_by(Sched) %>% summarize(n=n()) %>% ungroup()

schedule_names<-c('Chemicals, Oil, and Paints', 'Earths, Earthenware, and Glassware', 'Metals and Manufactures of', 'Wood and Manufactures of', 'Sugar, Molasses, and Manufactures of', 'Tobacco and Manufactures of', 'Agricultural Products and Provisions', 'Spirits, Wines, and other Beverages','Cotton Manufactures','Flax, Hemp, Jute, and Manufactures of','Wool and Manufactures of', 'Silk Manufactures','Manufactures of Rayon or Other Synthetic Textile', 'Papers and Books', 'Sundries')

abbr <- c('Chemicals', 'Glass', 'Metals', 'Wood', 'Sugar', 'Tobacco', 'Ag', 'Spirits','Cotton','Flax','Wool', 'Silk','Rayon', 'Paper', 'Sundries')

data_schedules<-data.frame(para_count,abbr,schedule_names)
names(data_schedules)<-c('Schedule','Items','Abbreviation','Title')

s <- kable(data_schedules, align= 'ccl',booktabs=T) %>%
  kable_styling(full_width = F, latex_options = "hold_position", font_size = 5)
add_header_above(s, c("Schedule Titles"=4),font_size = 9)

```


### Table 6
```{r, include=F}

change_BG <-
  data_set %>%
  filter(
    is.na(Specific_SH) != is.na(Specific_1946_after) | 
      is.na(Ad_Valorem_SH) != is.na(Ad_Valorem_1946_after) | Units_SH != Units_1946_after 
  ) %>%
  mutate(BG = 1,unit_ch = ifelse(Units_SH != Units_1946_after, 1, 0))

change_G <-
  data_set %>%
  filter(
    is.na(Specific_1946_after) != is.na(Specific_Geneva) | 
      is.na(Ad_Valorem_1946_after) != is.na(Ad_Valorem_Geneva) | Units_1946_after != Units_Geneva 
  ) %>%
  mutate(G = 1,unit_ch = ifelse(Units_1946_after != Units_Geneva, 1, 0))

change_A <-
  data_set %>%
  filter(
    is.na(Specific_Annecy) != is.na(Specific_Geneva) | 
      is.na(Ad_Valorem_Annecy) != is.na(Ad_Valorem_Geneva) | Units_Annecy != Units_Geneva
  ) %>%
  mutate(A = 1,unit_ch = ifelse(Units_Annecy != Units_Geneva, 1, 0))

change_T <-
  data_set %>%
  filter(
    is.na(Specific_Annecy) != is.na(Specific_Torquay) | 
      is.na(Ad_Valorem_Annecy) != is.na(Ad_Valorem_Torquay) | Units_Annecy != Units_Torquay
  ) %>%
  mutate(T = 1,unit_ch = ifelse(Units_Annecy != Units_Torquay, 1, 0))

change_GA <-
  data_set %>%
  filter(
    is.na(Specific_Geneva56_A) != is.na(Specific_Torquay) | 
      is.na(Ad_Valorem_Geneva56_A) != is.na(Ad_Valorem_Torquay) | Units_Geneva56_A != Units_Torquay
  ) %>%
  mutate(GA = 1,unit_ch = ifelse(Units_Torquay != Units_Geneva56_A, 1, 0))

change_GB <-
  data_set %>%
  filter(
    is.na(Specific_Geneva56_A) != is.na(Specific_Geneva56_B) | 
      is.na(Ad_Valorem_Geneva56_A) != is.na(Ad_Valorem_Geneva56_B) | Units_Geneva56_A != Units_Geneva56_B
  ) %>%
  mutate(GB = 1,unit_ch = ifelse(Units_Geneva56_A != Units_Geneva56_B, 1, 0))

change_GC <-
  data_set %>%
  filter(
    is.na(Specific_Geneva56_C) != is.na(Specific_Geneva56_B) | 
      is.na(Ad_Valorem_Geneva56_C) != is.na(Ad_Valorem_Geneva56_B) | Units_Geneva56_C != Units_Geneva56_B
  ) %>%
  mutate(GC = 1,unit_ch = ifelse(Units_Geneva56_B != Units_Geneva56_C, 1, 0))

change_DA <-
  data_set %>%
  filter(
    is.na(Specific_Dillon_A) != is.na(Specific_Geneva56_C) | 
      is.na(Ad_Valorem_Dillon_A) != is.na(Ad_Valorem_Geneva56_C) | Units_Dillon_A != Units_Geneva56_C
  ) %>%
  mutate(DA = 1,unit_ch = ifelse(Units_Geneva56_C != Units_Dillon_A, 1, 0))

change_DB <-
  data_set %>%
  filter(is.na(Specific_Dillon_A) != is.na(Specific_Dillon_B) | 
      is.na(Ad_Valorem_Dillon_A) != is.na(Ad_Valorem_Dillon_B) | Units_Dillon_A != Units_Dillon_B
  ) %>%
  mutate(DB = 1,unit_ch = ifelse(Units_Dillon_B != Units_Dillon_B, 1, 0))

mergeCols <- colnames(data_set)
mC <- mergeCols[mergeCols != "unit_ch"]

changes <- select(change_BG,-unit_ch) %>% 
  full_join(select(change_G,-unit_ch),  by = mergeCols)%>%
  full_join(select(change_A,-unit_ch),  by = mergeCols) %>%
  full_join(select(change_T,-unit_ch),  by = mergeCols) %>% 
  full_join(select(change_GA,-unit_ch),  by = mergeCols) %>%
  full_join(select(change_GB,-unit_ch),  by = mergeCols) %>% 
  full_join(select(change_GC,-unit_ch),  by = mergeCols) %>%
  full_join(select(change_DA,-unit_ch),  by = mergeCols) %>%
  full_join(select(change_DB,-unit_ch),  by = mergeCols) %>% arrange(id)
```


```{r table_6, echo=FALSE, warning=FALSE, message=F}

changes2 <- changes %>% select(id,Sched,Paragraph,Product,Interval,starts_with("Specific"))
nochanges <- rows_delete(specific, changes2, by = "id")

sum_specific_nC_pc <- nochanges %>%
  group_by(Sched) %>%
  summarize(SH_mean = mean(Specific_SH, na.rm=TRUE),
            DB_mean = mean(Specific_Dillon_B, na.rm=TRUE),
            mean_chg = ((SH_mean - DB_mean)/SH_mean)*100,
            SH_med = median(Specific_SH, na.rm=TRUE),
            DB_med = median(Specific_Dillon_B, na.rm=TRUE),
            med_chg = ((SH_med - DB_med)/SH_med)*100,
            n_specific = sum(!is.na(Specific_SH)),
            n = n()) %>%
  ungroup() # ungrouping variable is a good habit to prevent errors

tab6 <- kbl(sum_specific_nC_pc,digits=2,booktabs=T, caption = "Reduction in Specific Tariffs by Schedule") %>% kable_styling(position = "center", latex_options = "hold_position",font_size=4)

# Add header above means/medians for clarity
add_header_above(tab6, c(" " = 1, "Mean Comparisons "=3, "Median Comparisons"=3, "Observations" = 2),font_size = 4)
```


### Table 7
```{r table_7, message =F}
changesav <- changes %>% select(id,Sched,Paragraph,Product,Interval,starts_with("Ad_Valorem"))
nochanges_av <- rows_delete(ad_valorem, changesav, by = "id")

sum_av_nC_pc <- nochanges_av %>%
  group_by(Sched) %>%
  summarize(SH_mean = mean(Ad_Valorem_SH, na.rm=TRUE),
            DB_mean = mean(Ad_Valorem_Dillon_B, na.rm=TRUE),
            mean_chg = ((SH_mean - DB_mean)/SH_mean)*100,
            SH_med = median(Ad_Valorem_SH, na.rm=TRUE),
            DB_med = median(Ad_Valorem_Dillon_B, na.rm=TRUE),
            med_chg = ((SH_med - DB_med)/SH_med)*100,
            n_av = sum(!is.na(Ad_Valorem_SH)),
            n = n()) %>%
  ungroup() # ungrouping variable is a good habit to prevent errors

tab7 <- kbl(sum_av_nC_pc,digits=2,booktabs=T, caption = "Reduction in Ad Valorem Tariffs by Schedule") %>%
  kable_styling(position = "center",font_size=4)

# Add header above means/medians for clarity
add_header_above(tab7, c(" " = 1, "Mean Comparisons "=3, "Median Comparisons"=3, "Observations" = 2),font_size = 4)
```


# Notable Findings

### Importance of pre-GATT negotiations


### Some lines see tariff increases


### Very few changes between *ad valorem* and specific



# Next steps



## What's next?

- Concordances
  - Smoot-Hawley to TSUS
  - TSUS to HS
  - Smoot-Hawley to 1930's import classification system
- Add final three GATT rounds
  - Integrate "Free Lists" (Schedule 15, ~200 items by Torquay)
- Import volume and value data
  - *Ad valorem* equivalents
  - Trade weighting
  - Terms-of-trade analysis
- Other countries' schedules