---
title: "results"
author: "Kristy Buzard"
date: 'last edited `r Sys.Date()` '
output: pdf_document
---
```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(qwraps2)
library(kableExtra)
library(data.table)
library(scales)
```

<!-- [add roadmap for this section] -->

## Summarizing U.S. tariff liberalization: 1930-1964

### First look: item level

Our dataset covers the Smoot-Hawley round up through the Dillon (fifth) round of GATT negotiations. We begin with the `r dplyr::n_distinct(data_set$Paragraph)` paragraphs in the Smoot-Hawley Act and then add observations to our dataset whenever more detail is required.^[Note that there are more items than paragraphs because some paragraphs contain multiple tariff entries.] As new negotiations take place, products listed as one item---and thus have a single tariff rate---are often split into multiple items with different tariff rates. It is this phenomenon that leads to virtually all the added observations in our final dataset, which has `r nrow(data_set)` observations. We refer to an observation in our dataset as an `item.'

When the schedule for a round does not have information for a given item, we assume that the previous tariff rate is in effect. That is, we copy forward an item's tariff rate until a new rate appears as we move chronologically through the schedules. Our dataset is therefore a balanced panel.

This panel has ten variables that correspond to each of the available schedules: Smoot-Hawley (1930), the tariffs in place just before GATT negotiations began (1946), Geneva 1947, Annecy (1949), Torquay (1951), three phases for Geneva 1956 (1956, 1957, 1958) and two phases for Dillon (1963, 1964).

```{r echo=F}
same <- function(b,e){
  beg <- shortnames %>% select(ends_with(b))
  end <- shortnames %>% select(ends_with(e))
  name <- substitute(e)
  z <- as.name(paste0("same",name))
  assign(deparse(substitute(z)), shortnames %>%
           filter(((is.na(end[,3]) == T & is.na(beg[,3]) == T) | beg[,3] == end[,3])
                & ((is.na(end[,2]) == T & is.na(beg[,2]) == T) | beg[,2] == end[,2])
                & ((is.na(end[,1]) == T & is.na(beg[,1]) == T) | beg[,1] == end[,1]) )
         , envir=.GlobalEnv)
}  

# below we get the sets of lines for each round compared to SH
same("SH","DB")
same("SH","BG")
same("SH","Ge")
same("SH","An")
same("SH","To")
same("SH","GA")
same("SH","GB")
same("SH","GC")
same("SH","DA")
same("SH","DB")

nr = nrow(shortnames)
num_sameSH <- c(nr,nrow(sameBG),nrow(sameGe),nrow(sameAn),
                nrow(sameTo),nrow(sameGA),nrow(sameGB),
                nrow(sameGC),nrow(sameDA),nrow(sameDB))

same2 <- function(b,e){
  beg <- shortnames %>% select(ends_with(b))
  end <- shortnames %>% select(ends_with(e))
  name1 <- substitute(b)
  name2 <- substitute(e)
  z <- as.name(paste0("same",name1,"_",name2))
  assign(deparse(substitute(z)), shortnames %>%
           filter(((is.na(end[,3]) == T & is.na(beg[,3]) == T) | beg[,3] == end[,3])
                & ((is.na(end[,2]) == T & is.na(beg[,2]) == T) | beg[,2] == end[,2])
                & ((is.na(end[,1]) == T & is.na(beg[,1]) == T) | beg[,1] == end[,1]) )
         , envir=.GlobalEnv)
}  
# all the lines that are exactly the same in Smoot Hawley and Before Geneva
same2("SH","BG")

# below we get the sets of lines for each round compared to the previou round
same2("BG","Ge")
same2("Ge","An")
same2("An","To")
same2("To","GA")
same2("GA","GB")
same2("GB","GC")
same2("GC","DA")
same2("DA","DB")

year <- c(1930,1946,1947,1949,1951,1956,1957,1958,1963,1964)
num_nego <- c(0,nr-nrow(sameSH_BG),nr-nrow(sameBG_Ge),nr-nrow(sameGe_An),nr-nrow(sameAn_To),
            nr-nrow(sameTo_GA),nr-nrow(sameGA_GB),nr-nrow(sameGB_GC),
            nr-nrow(sameGC_DA),nr-nrow(sameDA_DB))
plot1_data <- data.frame(year,num_nego,num_sameSH)
```

```{r ibi, fig.cap = "Tariff Reductions by Number of Items", echo=F}
# Plot
ggplot() + 
  geom_line(data = plot1_data, aes(x=year, y= num_sameSH), color = "blue") +
  annotate("text", label = "No. of items still at Smoot-Hawley level", x = 1948, y = 2700, size = 5, colour = "blue") +
  geom_point(data = plot1_data, aes(x=year, y= num_nego), color = "red") +
  annotate("text", label = "No. of items reduced since previous round", x = 1952, y = 50, size = 5, colour = "red") +
  labs(x="Year", y= "Number of Items") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_x_continuous(breaks = seq(1930, 1965, by = 5)) +
  scale_y_continuous(breaks = seq(0, 3000, by = 500))
```

The two related, but distinct measures shown in Figure \@ref(fig:ibi) provide a first glimpse into the breadth of negotiations.^[Note that the 1946---just pre-Geneva 1947---embodied bilateral negotiations with nearly twenty countries, the results of which were extended on an MFN basis.] The red dots show how many items saw new commitments in each round. Geneva 1947 saw the largest number of items with commitments of any round: `r nr-nrow(sameBG_Ge)`. The Annecy round saw the fewest new commitments at `r nr-nrow(sameGe_An)`; this is in line with expectations since the U.S. only negotiated with new contracting parties in this round.

The blue line connects the dots indicating the number of items that had no negotiated change from Smoot Hawley at each point in our data. For instance, between Smoot Hawley and the start of the Geneva talks in 1946, a little more than one-third of the items in our dataset had some negotiated commitment.^[Note that this does not imply that negotiators made 1000 commitments, as many items were split after 1946.] In the Geneva 1947 round, just under one-third of the items received commitments for the first time. By the end of the Dillon Round phase-in, only `r nrow(sameDB)` items remained with no negotiated commitment.

<!-- [ADD IF TIME: HOW MANY ITEMS SAW NEW COMMITMENTS IN EACH ROUND (THIS WOULD INCLUDE DUPLICATES)] -->

<!-- [ADD IF TIME: Which items from Smoot Hawley don't see any action? THESE ARE IN sameDB] -->


### From Smoot-Hawley to Dillon: what was accomplished?

```{r echo=FALSE, warning=FALSE}
av_med <- data.frame(ad_valorem[,6:15] %>% sapply(median, na.rm=TRUE))
av_mean <- data.frame(ad_valorem[,6:15] %>% sapply(mean, na.rm=TRUE))
av_count <- data.frame(ad_valorem[,6:15] %>% sapply(function(x) sum(!is.na(x))))
av_min <- data.frame(ad_valorem[,6:15] %>% sapply(min, na.rm=TRUE))
av_max <- data.frame(ad_valorem[,6:15] %>% sapply(max, na.rm=TRUE))
av_25p <- data.frame(ad_valorem[,6:15] %>% sapply(function(x) quantile(x,probs=0.25,na.rm = TRUE)))
av_75p <- data.frame(ad_valorem[,6:15] %>% sapply(function(x) quantile(x,probs=0.75,na.rm = TRUE)))

sp_mean <- data.frame(specific[,6:15] %>% sapply(mean, na.rm=TRUE))
sp_med <- data.frame(specific[,6:15] %>% sapply(median, na.rm=TRUE))
sp_count <- data.frame(specific[,6:15] %>% sapply(function(x) sum(!is.na(x))))
sp_min <- data.frame(specific[,6:15] %>% sapply(min, na.rm=TRUE))
sp_max <- data.frame(specific[,6:15] %>% sapply(max, na.rm=TRUE))
sp_25p <- data.frame(specific[,6:15] %>% sapply(function(x) quantile(x,probs=0.25,na.rm = TRUE)))
sp_75p <- data.frame(specific[,6:15] %>% sapply(function(x) quantile(x,probs=0.75,na.rm = TRUE)))
```

We find that from the Smoot-Hawley tariffs in 1930 to the last phase in of the Dillon Round in 1964, both specific and *ad valorem* tariffs were cut roughly in half. The mean *ad valorem* tariff binding decreases from `r round(av_mean["Ad_Valorem_SH",],1)`% in 1930 to `r round(av_mean["Ad_Valorem_Dillon_B",],1)`% in 1964. The medians are quite close to the means, dropping from `r round(av_med["Ad_Valorem_SH",],0)`% to `r round(av_med["Ad_Valorem_Dillon_B",],0)`%. Roughly two-thirds of the items in our dataset have an *ad valorem* tariff.

Things look a bit different for specific tariffs. The mean specific tariff binding decreases from `r round(sp_mean["Specific_SH",],0)` cents in 1930 to `r round(sp_mean["Specific_Dillon_B",],0)` cents in 1964, while the median bindings are much smaller, dropping from `r round(sp_med["Specific_SH",],2)` cents to `r round(sp_med["Specific_Dillon_B",],2)` cents. Specific tariffs are present for roughly half of the items in our data. Note that summary statistics throughout the paper for specific tariffs are *not* trade weighted; we are in the process of acquiring the trade data required to both trade weight summary statistics and compute *ad valorem* equivalents.

The number of items that have both *ad valorem* and specific components to their tariff in a given round varies from 475 to 505, or a little more than 15%.^[See Section \@ref(sched) below for more details.] Following the categorization in Teti (2000), we refer to these items as *compound*. These items are included in the previous statistics. Also included are *mixed* tariffs, that are specified as having either an *ad valorem* or specific rate, depending on which is higher. Roughly 10% of our items have some *mixed component*. And about 2% of the items are *technical* in nature, that is, being defined in non-standard units like the proportion of content that meets some criteria. We do not analyze these categories separately because, insofar as possible, we have integrated them with the basic *ad valorem* and specific types.

### Heterogeneity over time: Round-by-Round liberalization

In terms of the time pattern of liberalization, we can confirm a well-known stylized fact: the largest proportional cuts took place between the Smoot-Hawley legislation and the first round of the GATT in Geneva in 1947: average tariffs, both specific and *ad valorem*, fell by about 30% (see the following two tables).

```{r sp-rd}
sp_mean <- specific[,6:15] %>% sapply(mean, na.rm=TRUE)
sp_med <- specific[,6:15] %>% sapply(median, na.rm=TRUE)
av_mean <- ad_valorem[,6:15] %>% sapply(mean, na.rm=TRUE)
av_med <- ad_valorem[,6:15] %>% sapply(median, na.rm=TRUE)

sp_mt = cbind(data.frame(sp_mean),c(0,-diff(sp_mean)/data.frame(sp_mean)[-nrow(data.frame(sp_mean)),]*100),data.frame(sp_med),c(0,-diff(sp_med)/data.frame(sp_med)[-nrow(data.frame(sp_med)),]*100))

names(sp_mt)<-c('Mean','% decrease','Median','% decrease')
rownames(sp_mt) <- c("Smoot Hawley", "1946","Geneva", "Annecy","Torquay","GenevaA","GenevaB","GenevaC","DillonA","DillonB")
kbl(sp_mt,digits=2,booktabs=T, caption = "Decrease in Specific Tariffs by Round") %>% 
  kable_styling(position = "center", latex_options = "hold_position")
```

```{r av-rd}
av_mt = cbind(data.frame(av_mean),c(0,-diff(av_mean)/data.frame(av_mean)[-nrow(data.frame(av_mean)),]*100),data.frame(av_med),c(0,-diff(av_med)/data.frame(av_med)[-nrow(data.frame(av_med)),]*100))

names(av_mt)<-c('Mean','% decrease','Median','% decrease')
rownames(av_mt) <- c("Smoot Hawley", "1946","Geneva", "Annecy","Torquay","GenevaA","GenevaB","GenevaC","DillonA","DillonB")
kbl(av_mt,digits=2,booktabs=T, caption = "Decrease in Ad Valorem Tariffs by Round") %>%
  kable_styling(position = "center", latex_options = "hold_position")
```

What will be surprising to many is the the United States was *very* active in negotiating bilateral agreements between 1934 and the start of the GATT negotiations in Geneva in 1947. In terms of *ad valorem* tariffs, a little less than half of the 30% reduction between Smoot Hawley and Geneva 1947 is directly attributable to the bilateral trade agreements because the U.S. converted all its bilateral concessions into MFN tariffs.^[Only a few pre-existing agreements, such as with Cuba and Philippines, did not have their commitments extended in this way.] Taking into account the 1946 tariff levels, the reduction in *ad valorem* tariffs we find is very close to the figure reported in Irwin (2017 p. 485) of 21%.

The pattern of reductions after Geneva 1947 is somewhat uneven. From Geneva 1947 to Annecy, the reduction in *ad valorem* tariffs was much smaller at about 3% for the mean and 20% for the medians. Reductions in specific tariffs were about half the size of the reductions in *ad valorem* tariffs. Annecy saw small tariff cuts on average in large part because this round was focused on accession: countries like the United States that had joined in Geneva only negotiated with those countries that were newly acceding in the Annecy Round.

In percentage terms, the Torquay Round saw cuts about two-thirds of the size of those in Geneva 1947. Geneva 1956 saw percentage cuts closer to those in Annecy, and for the first time these were phased in. The cuts in tariff averages in the Dillon round were about 5% for specific tariffs and 10% for *ad valorem*, while the median for both types of tariffs decreased by about 15%.


```{r sp-sm-db}
sp_table = cbind(data.frame(sp_min),data.frame(sp_25p),data.frame(sp_mean),data.frame(sp_med),data.frame(sp_75p),data.frame(sp_max),data.frame(sp_count))
names(sp_table)<-c('Min','1st Quartile','Mean','Median','3rd Quartile','Max','N')
rownames(sp_table) <- c("Smoot Hawley", "1946","Geneva", "Annecy","Torquay","GenevaA","GenevaB","GenevaC","DillonA","DillonB")
kbl(sp_table,digits=2,booktabs=T,caption = "Summary Statistics for Specific Tariffs by Round") %>%
  kable_styling(position = "center", latex_options = "hold_position")
```

```{r av-sm-db}
adval_table = cbind(data.frame(av_min),data.frame(av_25p),data.frame(av_mean),data.frame(av_med),data.frame(av_75p),data.frame(av_max),data.frame(av_count))
names(adval_table)<-c('Min','1st Quartile','Mean','Median','3rd Quartile','Max','N')
rownames(adval_table) <- c("Smoot Hawley","1946", "Geneva", "Annecy","Torquay","GenevaA","GenevaB","GenevaC","DillonA","DillonB")
kbl(adval_table,digits=2,booktabs=T,caption = "Summary Statistics of Ad Valorem Tariffs by Round") %>%
  kable_styling(position = "center", latex_options = "hold_position")
```


### Heterogeneity across products: Schedule-by-Schedule liberalization {#sched}

```{r}
change_BG <-
  data_set %>%
  filter(
    is.na(Specific_SH) != is.na(Specific_1946_after) | 
      is.na(Ad_Valorem_SH) != is.na(Ad_Valorem_1946_after) | Units_SH != Units_1946_after 
  ) %>%
  mutate(BG = 1,unit_ch = ifelse(Units_SH != Units_1946_after, 1, 0))
         
change_G <-
  data_set %>%
  filter(
    is.na(Specific_1946_after) != is.na(Specific_Geneva) | 
      is.na(Ad_Valorem_1946_after) != is.na(Ad_Valorem_Geneva) | Units_1946_after != Units_Geneva 
  ) %>%
  mutate(G = 1,unit_ch = ifelse(Units_1946_after != Units_Geneva, 1, 0))
         
change_A <-
  data_set %>%
  filter(
    is.na(Specific_Annecy) != is.na(Specific_Geneva) | 
      is.na(Ad_Valorem_Annecy) != is.na(Ad_Valorem_Geneva) | Units_Annecy != Units_Geneva
  ) %>%
  mutate(A = 1,unit_ch = ifelse(Units_Annecy != Units_Geneva, 1, 0))

change_T <-
  data_set %>%
  filter(
    is.na(Specific_Annecy) != is.na(Specific_Torquay) | 
      is.na(Ad_Valorem_Annecy) != is.na(Ad_Valorem_Torquay) | Units_Annecy != Units_Torquay
  ) %>%
  mutate(T = 1,unit_ch = ifelse(Units_Annecy != Units_Torquay, 1, 0))

change_GA <-
  data_set %>%
  filter(
    is.na(Specific_Geneva56_A) != is.na(Specific_Torquay) | 
      is.na(Ad_Valorem_Geneva56_A) != is.na(Ad_Valorem_Torquay) | Units_Geneva56_A != Units_Torquay
  ) %>%
  mutate(GA = 1,unit_ch = ifelse(Units_Torquay != Units_Geneva56_A, 1, 0))

change_GB <-
  data_set %>%
  filter(
    is.na(Specific_Geneva56_A) != is.na(Specific_Geneva56_B) | 
      is.na(Ad_Valorem_Geneva56_A) != is.na(Ad_Valorem_Geneva56_B) | Units_Geneva56_A != Units_Geneva56_B
  ) %>%
  mutate(GB = 1,unit_ch = ifelse(Units_Geneva56_A != Units_Geneva56_B, 1, 0))

change_GC <-
  data_set %>%
  filter(
    is.na(Specific_Geneva56_C) != is.na(Specific_Geneva56_B) | 
      is.na(Ad_Valorem_Geneva56_C) != is.na(Ad_Valorem_Geneva56_B) | Units_Geneva56_C != Units_Geneva56_B
  ) %>%
  mutate(GC = 1,unit_ch = ifelse(Units_Geneva56_B != Units_Geneva56_C, 1, 0))

change_DA <-
  data_set %>%
  filter(
    is.na(Specific_Dillon_A) != is.na(Specific_Geneva56_C) | 
      is.na(Ad_Valorem_Dillon_A) != is.na(Ad_Valorem_Geneva56_C) | Units_Dillon_A != Units_Geneva56_C
  ) %>%
  mutate(DA = 1,unit_ch = ifelse(Units_Geneva56_C != Units_Dillon_A, 1, 0))

change_DB <-
  data_set %>%
  filter(
    is.na(Specific_Dillon_A) != is.na(Specific_Dillon_B) | 
      is.na(Ad_Valorem_Dillon_A) != is.na(Ad_Valorem_Dillon_B) | Units_Dillon_A != Units_Dillon_B
  ) %>%
  mutate(DB = 1,unit_ch = ifelse(Units_Dillon_B != Units_Dillon_B, 1, 0))

mergeCols <- colnames(data_set)
mC <- mergeCols[mergeCols != "unit_ch"]
changes <- select(change_BG,-unit_ch) %>% 
  full_join(select(change_G,-unit_ch),  by = mergeCols)%>%
  full_join(select(change_A,-unit_ch),  by = mergeCols) %>%
  full_join(select(change_T,-unit_ch),  by = mergeCols) %>% 
  full_join(select(change_GA,-unit_ch),  by = mergeCols) %>%
  full_join(select(change_GB,-unit_ch),  by = mergeCols) %>% 
  full_join(select(change_GC,-unit_ch),  by = mergeCols) %>%
  full_join(select(change_DA,-unit_ch),  by = mergeCols) %>%
  full_join(select(change_DB,-unit_ch),  by = mergeCols) %>% arrange(id)
```

These headline numbers hide variation in magnitude and speed of liberalization across types of products. The data are grouped into 15 industry groups called *Schedules* in Smoot Hawley. They range in length from `r count(ad_valorem[which(ad_valorem$Sched == 6),])` items for Schedule 6 (Tobacco) to  `r count(ad_valorem[which(ad_valorem$Sched == 3),])` items for Schedule 3 (Metals).

```{r}
para_count <- data_set %>% group_by(Sched) %>% summarize(n=n()) %>% ungroup()

schedule_names<-c('Chemicals, Oil, and Paints', 'Earths, Earthenware, and Glassware', 'Metals and Manufactures of', 'Wood and Manufactures of', 'Sugar, Molasses, and Manufactures of', 'Tobacco and Manufactures of', 'Agricultural Products and Provisions', 'Spirits, Wines, and other Beverages','Cotton Manufactures','Flax, Hemp, Jute, and Manufactures of','Wool and Manufactures of', 'Silk Manufactures','Manufactures of Rayon or Other Synthetic Textile', 'Papers and Books', 'Sundries')

abbr <- c('Chemicals', 'Glass', 'Metals', 'Wood', 'Sugar', 'Tobacco', 'Ag', 'Spirits','Cotton','Flax','Wool', 'Silk','Rayon', 'Paper', 'Sundries')

data_schedules<-data.frame(para_count,abbr,schedule_names)
names(data_schedules)<-c('Schedule','Items','Abbreviation','Title')

s <- kable(data_schedules, align= 'ccl',booktabs=T) %>%
  kable_styling(full_width = F, latex_options = "hold_position")
add_header_above(s, c("Schedule Titles"=4),font_size = 12)
```

Note that 98 items switch between *ad valorem* and *specific* at some point. We exclude these items from the analysis in this section because they can distort the summary statistics for small groups.

Recall that about 16% of items in our sample have compound tariffs, that is, both a specific and *ad valorem* component. Table \@ref(tab:comp-sc) shows the variation across schedules in the rate of compound tariffs. Four of the smaller schedules have no compound tariffs at all, while over 60% of items in Schedul 11 (Wool) have compound tariffs. The schedule with the largest number of compound tariffs is Schedule 3 (Metals), which has 124 items with compound tariffs. 

```{r comp-sc}
both <- data_set %>% mutate(bothSH=!is.na(data_set[,"Specific_SH"])&!is.na(data_set[,"Ad_Valorem_SH"]),bothGeneva=!is.na(data_set[,"Specific_Geneva"])&!is.na(data_set[,"Ad_Valorem_Geneva"]),bothAnnecy=!is.na(data_set[,"Specific_Annecy"])&!is.na(data_set[,"Ad_Valorem_Annecy"]),bothTorquay=!is.na(data_set[,"Specific_Torquay"])&!is.na(data_set[,"Ad_Valorem_Torquay"]),bothGeneva56A=!is.na(data_set[,"Specific_Geneva56_A"])&!is.na(data_set[,"Ad_Valorem_Geneva56_A"]),bothGeneva56B=!is.na(data_set[,"Specific_Geneva56_B"])&!is.na(data_set[,"Ad_Valorem_Geneva56_B"]),bothGeneva56C=!is.na(data_set[,"Specific_Geneva56_C"])&!is.na(data_set[,"Ad_Valorem_Geneva56_C"]),bothDillonA=!is.na(data_set[,"Specific_Dillon_A"])&!is.na(data_set[,"Ad_Valorem_Dillon_A"]),bothDillonB=!is.na(data_set[,"Specific_Dillon_B"])&!is.na(data_set[,"Ad_Valorem_Dillon_B"]))

Both_group_by_sched<-table(both$Sched,both$bothGeneva)
Both_group_by_sched<-as.data.frame.matrix(Both_group_by_sched)

names(Both_group_by_sched)[1] <- "Not_compound_tariff"
names(Both_group_by_sched)[2] <- "Compound_tariff"

ratio<-label_percent()(Both_group_by_sched$Compound_tariff/(Both_group_by_sched$Not_compound_tariff+Both_group_by_sched$Compound_tariff))
Both_group_by_sched$schedule_names<-abbr

Both_group_by_sched$ratio<-ratio
Both_group_by_sched<-Both_group_by_sched[,c(3,1,2,4)]
kbl(Both_group_by_sched,col.names = c("Schedule","Not compound", "Compound", "Proportion compound"),digits=1,align= 'lccc',booktabs=T, caption = "Proportion of Compound Tariffs by Schedule ") %>% kable_styling(position = "center")
```

```{r sp-sc, echo=FALSE, warning=FALSE}
changes2 <- changes %>% select(id,Sched,Paragraph,Product,Interval,starts_with("Specific"))
nochanges <- rows_delete(specific, changes2, by = "id")

sum_specific_nC_pc <- nochanges %>%   
  group_by(Sched) %>%  
  summarize(SH_mean = mean(Specific_SH, na.rm=TRUE),
            DB_mean = mean(Specific_Dillon_B, na.rm=TRUE),
            mean_chg = ((SH_mean - DB_mean)/SH_mean)*100,
            SH_med = median(Specific_SH, na.rm=TRUE),
            DB_med = median(Specific_Dillon_B, na.rm=TRUE),
            med_chg = ((SH_med - DB_med)/SH_med)*100,
            n_specific = sum(!is.na(Specific_SH)),
            n = n()) %>%
  ungroup() # ungrouping variable is a good habit to prevent errors

kbl(sum_specific_nC_pc,digits=2,booktabs=T, caption = "Reduction in Specific Tariffs by Schedule") %>% kable_styling(position = "center", latex_options = "hold_position")
```

```{r av-sc}
changesav <- changes %>% select(id,Sched,Paragraph,Product,Interval,starts_with("Ad_Valorem"))
nochanges_av <- rows_delete(ad_valorem, changesav, by = "id")

sum_av_nC_pc <- nochanges_av %>%   
  group_by(Sched) %>% 
  summarize(SH_mean = mean(Ad_Valorem_SH, na.rm=TRUE),
            DB_mean = mean(Ad_Valorem_Dillon_B, na.rm=TRUE),
            mean_chg = ((SH_mean - DB_mean)/SH_mean)*100,
            SH_med = median(Ad_Valorem_SH, na.rm=TRUE),
            DB_med = median(Ad_Valorem_Dillon_B, na.rm=TRUE),
            med_chg = ((SH_med - DB_med)/SH_med)*100,
            n_av = sum(!is.na(Ad_Valorem_SH)),
            n = n()) %>%
  ungroup() # ungrouping variable is a good habit to prevent errors
kbl(sum_av_nC_pc,digits=2,booktabs=T, caption = "Reduction in Ad Valorem Tariffs by Schedule") %>%
  kable_styling(position = "center")
```
Similarly, there is a lot of variation in the proportions of specific and *ad valorem* tariffs across schedules. Schedule 8 (Spirits) has all but `r as.character(english(sum(is.na(specific[which(specific$Sched == 8),"Specific_Dillon_B"]))))` of its `r count(ad_valorem[which(ad_valorem$Sched == 8),])` items covered by a specific tariff. Sechedul 12 (Silk), on the other hand, has no specific tariffs. All 37 lines are covered by an *ad valorem* tariff only.

We begin with the caveat that the variation in types of units poses a particular challenge to interpretation here since units tend to be concentrated in particular schedules. That said, the largest decrease in mean specific tariffs is for Schedule 7 (Agriculture) at 70.5%, while the largest decrease in median specific tariffs is for Schedule 5 (Sugar) at 75%. In contrast, the `r as.character(english(sum(!is.na(ad_valorem[which(ad_valorem$Sched == 5),"Ad_Valorem_Dillon_B"]))))` items in Schedule 5 that have *ad valorem* tariffs see the smallest decrease across all schedules for mean *ad valorem* tariffs: from `r round(mean(ad_valorem[which(ad_valorem$Sched == 5),"Ad_Valorem_SH"]$Ad_Valorem_SH, na.rm = TRUE),1)`% to `r round(mean(ad_valorem[which(ad_valorem$Sched == 5),"Ad_Valorem_Dillon_B"]$Ad_Valorem_Dillon_B, na.rm = TRUE),1)`%. 

<!-- check the facts in this next paragraph by hand--we need to be aware if this changes -->

It is notable that all schedules have the mean *ad valorem* tariff decrease by less than the median *ad valorem* tariff decreases, implying a reduction in tariff spikes. For specific tariffs, two-thirds of the schedules show this pattern, while the Schedules for Chemicals, Wood, Sugar, Rayon and Paper have larger decreases for the medians.

```{r sp-sc-rd}
sum_specific_bySched <- nochanges %>%   
  group_by(Sched) %>%  
  summarize(SH = mean(Specific_SH, na.rm=TRUE),
            BG = mean(Specific_1946_after, na.rm=TRUE),
            Ge = mean(Specific_Geneva, na.rm=TRUE),
            An = mean(Specific_Annecy, na.rm=TRUE),
            To = mean(Specific_Torquay, na.rm=TRUE),
            GC = mean(Specific_Geneva56_C, na.rm=TRUE),
            DB = mean(Specific_Dillon_B, na.rm=TRUE),
            chBG = ((SH - BG)/SH)*100,
            chGe = ((BG - Ge)/BG)*100,
            chAn = ((Ge - An)/Ge)*100,
            chTo = ((An - To)/An)*100,
            chGC = ((To - GC)/To)*100,
            chDB = ((GC - DB)/GC)*100,
            ) %>%
  ungroup() # ungrouping variable is a good habit to prevent errors

kbl(sum_specific_bySched,digits=1,booktabs=T, caption = "Mean Specific Tariffs by Schedule and Round") %>%
  kable_styling(position = "center", latex_options = "hold_position")
```

```{r av-sc-rd}
sum_av_bySched_nC <- nochanges_av %>%   
  group_by(Sched) %>%  
  summarize(SH = mean(Ad_Valorem_SH, na.rm=TRUE),
            BG = mean(Ad_Valorem_1946_after, na.rm=TRUE),
            Ge = mean(Ad_Valorem_Geneva, na.rm=TRUE),
            An = mean(Ad_Valorem_Annecy, na.rm=TRUE),
            To = mean(Ad_Valorem_Torquay, na.rm=TRUE),
            GC = mean(Ad_Valorem_Geneva56_C, na.rm=TRUE),
            DB = mean(Ad_Valorem_Dillon_B, na.rm=TRUE),
            chBG = ((SH - BG)/SH)*100,
            chGe = ((BG - Ge)/BG)*100,
            chAn = ((Ge - An)/Ge)*100,
            chTo = ((An - To)/An)*100,
            chGC = ((To - GC)/To)*100,
            chDB = ((GC - DB)/GC)*100,
            ) %>%
  ungroup() # ungrouping variable is a good habit to prevent errors

kbl(sum_av_bySched_nC,digits=1,booktabs=T, caption = "Mean Ad Valorem by Schedule and Round") %>%
  kable_styling(position = "center", latex_options = "hold_position")
```

Tables \@ref(tab:sp-sc-rd) and \@ref(tab:av-sc-rd) show the round-by-round changes in specific and *ad valorem* tariffs for each schedule. "BG" denotes the change from Smoot Hawley to 1946 just before the Geneva Round, "Ge" from 1946 to Geneva, etc. "GC" is the change from Torquay to the third ("C") and final phase in of Geneva 1956; similarly, "DB" is the change from Geneva "C" to Dillon "B," the second and final phase in stage for Dillon). 

<!-- -   Any obvious patterns to which items have largest/smallest cuts? -->

<!-- -   Round by round graphs for specific, interesting items -->

## Notable Findings

<!-- ### Concertina-style liberalization -->

<!-- The literature classifies patterns of liberalization into two main types: uniform radial reduction and concertina (e.g. Neary 1998). A radial reduction takes place when all tariffs are reduced by the same proportion. In contrast, under a concertina reform, the highest tariff rates are reduced more than proportionately. Hoda (2001) finds that concertina reforms have been most popular recently, and we find evidence that the U.S. tariff reductions under the early rounds of the GATT also follow this pattern. -->

<!-- ```{r sp-dec} -->
<!-- library(stats) -->
<!-- Sp_chg_BG<-(shortnames$Sp_SH-shortnames$Sp_BG)*100/shortnames$Sp_SH -->
<!-- Sp_chg_BG_q<-quantile(Sp_chg_BG, probs = seq(0.1, 1, 0.1), na.rm = TRUE,) -->

<!-- Sp_chg_Ge<-(shortnames$Sp_BG-shortnames$Sp_Ge)*100/shortnames$Sp_BG -->
<!-- Sp_chg_Ge_q<-quantile(Sp_chg_Ge, probs = seq(0.1, 1, 0.1), na.rm = TRUE,) -->

<!-- Sp_chg_An<-(shortnames$Sp_Ge-shortnames$Sp_An)*100/shortnames$Sp_Ge -->
<!-- Sp_chg_An_q<-quantile(Sp_chg_An, probs = seq(0.1, 1, 0.1), na.rm = TRUE,) -->

<!-- Sp_chg_To<-(shortnames$Sp_An-shortnames$Sp_To)*100/shortnames$Sp_An -->
<!-- Sp_chg_To_q<-quantile(Sp_chg_To, probs = seq(0.1, 1, 0.1), na.rm = TRUE,) -->

<!-- Sp_chg_GC<-(shortnames$Sp_To-shortnames$Sp_GC)*100/shortnames$Sp_To -->
<!-- Sp_chg_GC_q<-quantile(Sp_chg_GC, probs = seq(0.1, 1, 0.1), na.rm = TRUE,) -->

<!-- Sp_chg_DB<-(shortnames$Sp_GC-shortnames$Sp_DB)*100/shortnames$Sp_GC -->
<!-- Sp_chg_DB_q<-quantile(Sp_chg_DB, probs = seq(0.1, 1, 0.1), na.rm = TRUE,) -->

<!-- Sp_Changes_quantile<-data.frame(cbind(Sp_chg_BG_q,Sp_chg_Ge_q,Sp_chg_An_q,Sp_chg_To_q,Sp_chg_GC_q,Sp_chg_DB_q)) -->
<!-- names(Sp_Changes_quantile)<-c('1946','Geneva','Annecy','Torquay','Geneva 56','Dillon') -->
<!-- rownames(Sp_Changes_quantile) <- c("10%","20%","30%","40%","50%","60%","70%","80%","90%","100%") -->

<!-- kbl(Sp_Changes_quantile,digits=1,booktabs=T, caption = "Percent Change in Specific Tariff by Decile") %>% kable_styling(position = "center") -->
<!-- ``` -->

<!-- ```{r av-dec} -->
<!-- AV_chg_BG<-(shortnames$AV_SH-shortnames$AV_BG)*100/shortnames$AV_SH -->
<!-- AV_chg_BG_q<-quantile(AV_chg_BG, probs = seq(0.1, 1, 0.1), na.rm = TRUE,) -->

<!-- AV_chg_Ge<-(shortnames$AV_BG-shortnames$AV_Ge)*100/shortnames$AV_BG -->
<!-- AV_chg_Ge_q<-quantile(AV_chg_Ge, probs = seq(0.1, 1, 0.1), na.rm = TRUE,) -->

<!-- AV_chg_An<-(shortnames$AV_Ge-shortnames$AV_An)*100/shortnames$AV_Ge -->
<!-- AV_chg_An_q<-quantile(AV_chg_An, probs = seq(0.1, 1, 0.1), na.rm = TRUE,) -->

<!-- AV_chg_To<-(shortnames$AV_An-shortnames$AV_To)*100/shortnames$AV_An -->
<!-- AV_chg_To_q<-quantile(AV_chg_To, probs = seq(0.1, 1, 0.1), na.rm = TRUE,) -->

<!-- AV_chg_GC<-(shortnames$AV_To-shortnames$AV_GC)*100/shortnames$AV_To -->
<!-- AV_chg_GC_q<-quantile(AV_chg_GC, probs = seq(0.1, 1, 0.1), na.rm = TRUE,) -->

<!-- AV_chg_DB<-(shortnames$AV_GC-shortnames$AV_DB)*100/shortnames$AV_GC -->
<!-- AV_chg_DB_q<-quantile(AV_chg_DB, probs = seq(0.1, 1, 0.1), na.rm = TRUE,) -->

<!-- AV_Changes_quantile<-data.frame(cbind(AV_chg_BG_q,AV_chg_Ge_q,AV_chg_An_q,AV_chg_To_q,AV_chg_GC_q,AV_chg_DB_q))  -->
<!-- names(AV_Changes_quantile)<-c('1946','Geneva','Annecy','Torquay','Geneva 56','Dillon') -->
<!-- rownames(AV_Changes_quantile) <- c("10%","20%","30%","40%","50%","60%","70%","80%","90%","100%") -->

<!-- kbl(AV_Changes_quantile,digits=1,booktabs=T, caption = "Percent Change in Mean Tariff by Decile") %>% kable_styling(position = "center") -->
<!-- ``` -->

<!-- Evidence: I created two tables to demonstrate the quantiles of percentage tax reduction (on product level) across every two rounds. In both ad valorem table and specific table we can see that 50% of lines are not changed, but for the highest 10%~20%, the reduction magnitude is significant, and this pattern remains across all rounds (I haven’t been able to track the highest reduction lines, but intuitively I think they should be different products, it is hard to imagine certain products keep having this much of tax reduction). This indicates that between every two round, the tax reduction is relatively centralized, or more focused on a certain range of products, instead of “uniform reduction” for all product listed. -->


### Importance of the RTAA in the pre-GATT years

Bown and Irwin (2017) present summary data from the USITC on the reduction in rates from Smoot-Hawley to just before Geneva and then to the Geneva rates. We cannot replicate this table until our trade volume and value data are complete. 
However, columns "chBG" (for change from Smoot Hawley to "before Geneva", i.e. 1946) and "chGe" (for change from "before Geneva" to Geneva (1947) in Tables \@ref(tab:sp-sc-rd) and \@ref(tab:av-sc-rd) provide the straight item-by-item averages across specific and *ad valorem* tariffs.

However the data is aggregated, it's clear that the bilateral negotiations between 1934 and 1946 made an important contribution to the statutory liberalization of this era. The headline numbers we calculate for mean tariffs in Tables \@ref(tab:sp-rd) and \@ref(tab:av-rd) are very much in line with the USITC calculations reported by Bown and Irwin (2017) of 21% pre-Geneva to post-Geneva. We find smaller figures of 20% and 13% for the mean reduction in specific and *ad-valorem* tariffs, as compared to the USITC calculation of 33%, so this is an area where we look forward to finding the effects of trade weighting the data.

### Some items see tariff increases

Although it is rare, we see 78 instances in which a tariff increases. Dillon is the round in which we observe the most tariff increases, with 39. 22 of the cases in the Dillon Round are for items that have both specific and *ad valorem* elements. Of those, 10 saw a high *ad valorem* tariff but a lower specific tariff moving from Geneva 1956 to Dillon. One saw both *ad valorem* and specific tariffs go up, and the rest kept the same specific tariff but increased the *ad valorem* tariff by up to 140% (this *ad valorem* increase from 25% to 60% was in Paragraph 1108 for wool woven fabrics). The rest appear to be straightforward increases. For instance, four items in Paragraph 371---bicycles---saw the *ad valorem* tariff increase from 15% to 22.5%. Four items in Paragraph 371 also had only specific tariffs and saw those increase by exactly 50%, e.g. from $2 per bicycle to $3 per bicycle.

There were 18 increases from Smoot-Hawly to 1946, with many of them in Schedule 3 (Metals). But we also see the specific tariff rate on frozen swordfish (Paragraph 717a) increase, from 0.125 cents to 0.1865 cents per ounce. The Annecy Round saw no tariff increases.

Although these increases are scattered across most of the Schedules, Glass, Metals, Agricultural Goods and Wool account for the majority of them.


### Switching from *ad valorem* to specific is rare

Given that inflation was reducing the incidence of specific tariffs (Irwin 2017) in this period, one might have expected specific tariffs to be converted into *ad valorem* tariffs to avoid the loss of protection. Another reason we might expect this conversion is simply that only 8% of U.S. tariffs were specific during the three decades staring in 1988 (Teti 2020).

However, during our period of 1930 to 1964, we saw only 98 lines make a switch between specific and *ad valorem*. Of these, 37 switched from specific to *ad valorem*, while 24 switched from *ad valorem* to specific. The rest either moved into or out of the compound category. These showed no clear pattern of moving toward or away from one type of tariff.

This means that the reduction in the prevalence of specific tariffs must come in the rounds that we have left to analyze. We expect that switching to the new tariff nomenclatures (TSUS and HTS) will play an important role in this process.


<!-- ## Analysis -->

<!-- -   What can we say about which / why items have ad valorem vs. specific? -->

<!--     -   Is there variation over time? -->

<!--     , but we intend to analyze them separately and provide explanation for which items get *ad valorem* protection, which items get specific tariffs, and which get both. We know that the incidence of specific tariffs declines with inflation; this is an important and under-studied aspect of liberalization that we intend to explore. -->

<!-- -   Which products `suffer' from large tariff cuts and which products continue to receive protection? -->

