---
title: "Coding Torquay"
author: "Ross Jestrab"
date: "6/12/2020"
output: html_document
---

## Overview
This document discusses the process of digitizing the Torquay pdf. It also provides the R code used to help with this process. The final clean excel file with the Torquay data is called "Torquay All Schedules" in "G:\\MAX-Filer\\Collab\\SOP-kbuzard-S13\\Admin\\torquay".

## Preparing Torquay pdf
The "torquay (black white)" pdf in the Admin folder ("G:\\MAX-Filer\\Collab\\SOP-kbuzard-S13\\Admin") is used in this section (Note: The "torquay (color)" pdf also in the Admin folder is the same document, but is in color and is not as clear.). This document has 2 pages of the Torquay document on each pdf page. This format is difficult for R to digitize. The ideal format is one page of the Torquay document on each pdf page. To accomplish this I do the following: 

1. Create a copy of the "torquay (black white)" pdf file so that the original file will not be affected. 
2. Use the Adobe Pro "Crop Pages" tool under "Edit PDF" to crop the left hand side of all the pages and save this file as new file (using "save as"). Repeat this process to crop the right hand side of all the pages and save this as another file (using "save as"). 
3. Print out these 2 new files using the "Fit" option under the "Page Sizing & Handling" heading to increase the size of the text. This makes the files easier to read. Once both files are printed, organize them such that the pages are in increasing order. 
4. Separate this file into the 15 different schedules and use a document scanner to scan each schedule separately. Make sure to use the highest quality settings available (such as 600 dpi, black and white) and to select letter size. 
5. Using Adobe Pro, each schedule document needs to be word searchable (also known as OCR) by selecting the "Edit PDF" option under the tools tab. Make sure to select "Make all the pages editable" option by clicking "Settings" under the "SCANNED DOCUMENTS" tab. Remember to save after it completes the process. It may take a few minutes for the OCR process to finish for larger documents. Each schedule pdf (called "Schedule 1", "Schedule 2", etc.) is located in its own folder in "G:\\MAX-Filer\\Collab\\SOP-kbuzard-S13\\Admin\\torquay\\Individual Torquay Schedules".
6. Each schedule pdf is separated into single pages by using the "Extract Pages" option under the "Page Thumbnails". Select the "Extract Pages as Separate Files Option". This step is necessary since the R program has a difficult time with multiple pages at a single time. If the "pdftools" package is updated later to deal with this then this step of saving the schedule pdfs as single pages can be skipped. The separate pages for each schedule are located within their respective schedule folder in "G:\\MAX-Filer\\Collab\\SOP-kbuzard-S13\\Admin\\torquay\\Individual Torquay Schedules". For example, the "Schedule 1 1" pdf is the first page of schedule 1, and the "Schedule 1 2" pdf is the second page of schedule 1 (both located in "G:\\MAX-Filer\\Collab\\SOP-kbuzard-S13\\Admin\\torquay\\Individual Torquay Schedules\\Torquay Schedule 1").
7. You are now ready to use R.

## Using R
R has the "pdftools" package which allows for the digitization of pdf documents. Below I have included the general simple program I wrote to extract the appropriate information. In each schedule folder in "G:\\MAX-Filer\\Collab\\SOP-kbuzard-S13\\Admin\\torquay\\Individual Torquay Schedules" the R code is located in the R file called "pdf reader 1930 one page at a time (Torquay)". After running the R code, each page is saved as a text file. For example, page 1 of schedule 1 is saved as "schedule_1_1" in "G:\\MAX-Filer\\Collab\\SOP-kbuzard-S13\\Admin\\torquay\\Individual Torquay Schedules\\Torquay Schedule 1". Page 2 of schedule 1 is saved as "schedule_1_2" in the same folder. The text files can be opened with the Notepad program on Windows (or TextEdit on Mac). Please note that after the R program is ran the user will still have to do more cleaning of the data by hand (e.g. checking for spelling errors, organizing information into the appropriate columns/rows, etc.), but this program will help with having to type everything in. This manual cleaning is discussed in the next section. 

The below R program saves all the text after the first "Duty". This program loops over all the pages in Schedule 1. To run the R code, copy and paste it into a new R script file. The working directory location will have to be changed with the "setwd()" command.

```{r, eval=FALSE}
library(pdftools)
library(tidyverse)

rm(list = ls())
  
#creating a loop over all the files. 23 is the last page of the first schedule.
for (i in 1:23) {
  
  #----
  page <- i
  str_extract_start <- "(?<=Duty\\s).*" #may need to use (?<=PARAGRAPH\\s) or someother word.
  #----

  setwd("~/Desktop/Professor Buzard/Torquay Schedule 1") #for my computer. 

  pdf_1930 <- paste("Schedule 1", page, sep = " ")
  pdf_1930 <- paste(pdf_1930, "pdf", sep = ".")
  txt_save <- paste("schedule_1", page, sep = "_")

  text <- pdf_text(pdf_1930) %>%
    str_squish() 
  text

  text <- pdf_text(pdf_1930) %>%
    str_squish() %>%
    str_extract(str_extract_start) %>%
    str_split("\\.\\s") %>% 
    unlist() %>%
    plyr::ldply(.id = NULL) %>%
    as.data.frame() %>%
    write_tsv(txt_save)
}
```

Sometimes the OCR process does not recognize the "Duty" word correctly. If this happens for certain pages then "str_extract_start" can be change to a different word such as "paragraph" and input the specific page number in "page". An example of this for page 1 of Schedule 1 is included below. Note the loop is removed for this to insure only the specific pages are addressed. Again, the working directory location will have to be changed with the "setwd()" command.

```{r, eval=FALSE}
library(pdftools)
library(tidyverse)

rm(list = ls())

#----
page <- 1
str_extract_start <- "(?<=Duty\\s).*" #may need to use (?<=paragraph\\s) or someother word.
#----

setwd("~/Desktop/Professor Buzard/Torquay Schedule 1") #for my computer. 

pdf_1930 <- paste("Schedule 1", page, sep = " ")
pdf_1930 <- paste(pdf_1930, "pdf", sep = ".")
txt_save <- paste("schedule_1", page, sep = "_")

text <- pdf_text(pdf_1930) %>%
  str_squish() 
text

text <- pdf_text(pdf_1930) %>%
  str_squish() %>%
  str_extract(str_extract_start) %>%
  str_split("\\.\\s") %>% 
  unlist() %>%
  plyr::ldply(.id = NULL) %>%
  as.data.frame() %>%  
  write_tsv(txt_save)
```

## Manual Cleaning
To edit the files created from the above R program I use TextEdit on a Mac but any basic text editor should work such as Notepad on Windows. The steps used are the following:

1. Take the first page text file of each schedule and use "save as" to create a file that will become the master text file for each schedule. For example, it is called "Torquay Schedule 1" for schedule 1 in "G:\\MAX-Filer\\Collab\\SOP-kbuzard-S13\\Admin\\torquay\\Individual Torquay Schedules\\Torquay Schedule 1".
2. Copy and paste the remaining text file pages for each schedule to each master text file. 
3. At the top of each master file I insert the column heading names I want to include and use a tab to separate each column since I want the file to follow a tab delimited format. A tab delimited format separates each column with a tab and each row with a return. By working with this format it is faster to clean the data than say an excel file. It also makes importing the data into excel later easy.

Column Headings | Descriptions of Columns
- | -
Paragraph | The corresponding Tariff Act of 1930 paragraph number from the Torquay document. If there are sub-paragraphs there are separated by periods.
Description | Follows the "Description of Products" from the Torquay document. A detailed explanation of how this is entered follows this table.
Specific | The specific tariff for the row from the Torquay document. If there is not a specific tariff for the row, then leave it blank.
Units | The corresponding units for the specific tariff in the Specific column from the Torquay document. Leave this blank if the row has no specific tariff. I create the following coding system below to help with later data analysis (see Units Formatting below). The user can either delete the corresponding unit texts and replace it with the coding system below at this step or the user can keep the text and use another program such as R to replace the text later. I use a combination of the preceding methods for entering the Units information.
AdValorem | The ad valorem tariff for the row from the Torquay document. If there is no ad valorem tariff for the row, then leave it blank. As an example, I enter 10% as 10.
Extra | Collects any additional information from the Torquay document that may be of interests for later analysis (such as more detailed unit information, minimum tariffs, etc.). I need to decide if I want minimum tariffs to be included here or as a separate row (later in the main excel file "Geneva, Annecy, Torquay All Schedules" in "G:\\MAX-Filer\\Collab\\SOP-kbuzard-S13\\Admin\\torquay" that combines all tariff schedules I make the minimum and maximum a separate row.)
Schedule | The schedule number from the Tariff Act of 1930, which follows from the Tariff Act of 1930 paragraph number from the Torquay document. 
---------------

### Description Column Details
Here I discuss how I enter the information in the Description column. Many of the rows in the Torquay document are connected to previous rows descriptions. For example, look at the first two product in paragraph 1 from Schedule 1. In the pdf document it reads as the following,

Acetic acid containing by weight of 
  acetic acid -
    Not over 65%
    Over 65%

What this implies is the following:

Acetic acid containing by weight of acetic acid: Not over 65%

Acetic acid containing by weight of acetic acid: Over 65%


The format I use for the Descriptions column is the following:

Acetic acid containing by weight of acetic acid: Not over 65%

Over 65%


I use this format since it is easier to go between the pdf and the digitize versions. Additionally, it helps to cut-down on additional typing since some of the overlapping descriptions can be quite long for certain paragraphs. This means if you are interested in reading the description of a product make sure to look at the preceding rows for the same paragraph. In the main excel file "Geneva, Annecy, Torquay All Schedules" in "G:\\MAX-Filer\\Collab\\SOP-kbuzard-S13\\Admin\\torquay" that combines all tariff schedules I am currenlty in the process of making each description independent of the previous one (i.e. use the format from the "What this implies is the following:" from above).

### Units Formatting
Below is a table of the numbering used in the Units columns and the corresponding meaning. Please add any additional units when necessary.

Number | Meaning
- | -
1 | Cents per pound.
2 | Cents per gallon.
3 | Dollars per pound.
4 | Cents per ounce.
5 | Dollars per ounce.
6 | Cents per square foot.
7 | Cents per ton.
8 | Dollars per ton.
9 | Dollars per gallon.
10 | Cents per cubic foot.
---------------

## Making a Master Excel File for Torquay Data
To create a master excel file I do the following:

1. Create a master Torquay text file that includes all the schedules by copying and pasting all the schedule infomration from each schdule that is located in each folder in "G:\\MAX-Filer\\Collab\\SOP-kbuzard-S13\\Admin\\torquay\\Individual Torquay Schedules". 
a. Save the Schedule 1 using "save as" as "Torquay All Schedules" (note that the text version of "Torquay All Schedules" is deleted since it is converted to an excel file in the below step).
b. Copy and paste all the other schedule information entered from above and save the file.
2. Open Excel and import the newly created master text file. 
a. Choose the tab delimited option and the tab option. All the data will now be imported into a basic excel sheet. Save this file as "Torquay All Schedules" in "G:\\MAX-Filer\\Collab\\SOP-kbuzard-S13\\Admin\\torquay".
b. This is a good time to go through the document and make sure everything is is the appropriate row and column. It is easy to hit the tab button too many (or too few) of times in the previous cleaning. Also make sure the descriptions are the correct format. Sometimes excel adds "-" between words and check a few column to the right to make sure no information ended up there. When entering the other rounds information it is a good time to check the previously entered tariff numbers.
3. Save the excel file as a csv file so that it can easily be used in multiple programs for analysis, such as R.
