---
title: "Main Data: Combining Geneva, Annecy, Torquay Schedules"
author: "Ross Jestrab"
date: "6/12/2020"
output: html_document
---

## Overview

This file covers the process for making the master excel file that covers Geneva, Annecy, and Torquay (called "Geneva, Annecy, Torquay All Schedules" in "G:\\MAX-Filer\\Collab\\SOP-kbuzard-S13\\Admin\\torquay"). The file has all the Torquay data entered. Currently the Geneva and Annecy data is being added to it.

The strategy used to combine this data is I first coded all the Torquay data (see "Coding Torquay" R Markdown and html file in same folder for additional details.), then I manually go back through the Geneva and Annecy pdf documents to add their data to create a master excel file. I find it easier to print out paper versions of the Geneva, Annecy, and Torquay files when inputting the data into the excel file. Once all this data is entered an R program can be used to fill in the blank cells (see "Cleaning Data with R" section below).

The first 3 GATT rounds follow the Tariff Act of 1930 paragraphs for organizing the products. Most of the descriptions line up across these GATT rounds, but there are a few cases where they jump from specific products to more general descriptions (e.g., paragraphs 2, 5, and 23). After reading the descriptions, I am mostly able to figure out which products the tariffs correspond to (yellow highlighted lines in the excel file need to be checked in more detail).

## Data Entry Status

Below is the current status of the data entry. Note: Row 1 has the variables, discussed in the next section.

Excel Rows | Status
- | -
2 to 441 | Mostly complete. The rows that are highlighted yellow need to be checked since at first pass it was not clear how to match the descriptions of products across GATT rounds.
442 to 522 | Not done. Need to enter Geneva and Annecy data.
523 to 827 | Mostly done. The rows that are highlighted yellow need to be checked since at first pass it was not clear how to match the descriptions of products across GATT rounds.
828 to 2281 | Not done. Need to enter Geneva and Annecy data. Very few random lines inputted, but should be checked.
Notes: 

1) After all data is entered some process needs to be implemented to check the data since it is very easy to get one line off during data entry process.

2) Lines 2 to 441, and 523 to 965 have their "Description" variable as independent. This means that you do not have to look to a previous line to fully understand what the description is (see "Description Column Details" section from "Coding Torquay" R markdown and html file in the same folder for more information on this). Lines 442 to 522, and 966 to 2281 need for their description to be independent. The way I was doing this was by copying and pasting the necessary information from the previous rows.

## Variables

The variables are located in the first row of the excel file.

Variable Name | Variable Description
------------- | --------------------
Schedule | The schedule number from the Tariff Act of 1930, which follows from the Tariff Act of 1930 paragraph number (see "Schedule Numbering" below).
Paragraph | The corresponding Tariff Act of 1930 paragraph number. If there are sub-paragraphs there are separated by periods. Some additional sub-groupings use a "_" (e.g., see row 866 in excel file), need to decide if this is the correct or best way to do this.
Description | Follows the "Description of Products" from the Torquay document, but is split into multiple rows when Geneva and/or Annecy split the products. This separating can be somewhat subjective if the descriptions do not exactly match across products.
Extra | Collects any additional information from the Geneva, Annecy, and Torquay files that may be of interests for later analysis (such as more detailed unit information, etc.). If a product has a minimum and maximum tariff I try to enter each as a separate row.
Specific_Geneva | The specific tariff for the row from the Geneva document. If there is not a specific tariff for the row, then leave it blank.
Units_Geneva | The corresponding units for the specific tariff in the Specific_Geneva column from the Geneva document. Leave this blank if the row has no specific tariff. I create the following coding system below to help with later data analysis (see "Units Formatting" below). The user can either delete the corresponding unit texts and replace it with the coding system below at this step or the user can keep the text and use another program such as R to replace the text later. I use a combination of the preceding methods for entering the Units information.
Ad_Valorem_Geneva | The ad valorem tariff for the row from the Geneva document. If there is no ad valorem tariff for the row, then leave it blank. As an example, I enter 10% as 10.
Specific_Annecy | The specific tariff for the row from the Annecy document. If there is not a specific tariff for the row, then leave it blank.
Units_Annecy | The corresponding units for the specific tariff in the Specific_Annecy column from the Annecy document. Leave this blank if the row has no specific tariff. I create the following coding system below to help with later data analysis (see "Units Formatting" below). The user can either delete the corresponding unit texts and replace it with the coding system below at this step or the user can keep the text and use another program such as R to replace the text later. I use a combination of the preceding methods for entering the Units information.
Ad_Valorem_Annecy | The ad valorem tariff for the row from the Annecy document. If there is no ad valorem tariff for the row, then leave it blank. As an example, I enter 10% as 10.
Specific_Torquay | The specific tariff for the row from the Torquay document. If there is not a specific tariff for the row, then leave it blank.
Units_Torquay | The corresponding units for the specific tariff in the Specific_Torquay column from the Torquay document. Leave this blank if the row has no specific tariff. I create the following coding system below to help with later data analysis (see "Units Formatting" below). The user can either delete the corresponding unit texts and replace it with the coding system below at this step or the user can keep the text and use another program such as R to replace the text later. I use a combination of the preceding methods for entering the Units information.
Ad_Valorem_Torquay | The ad valorem tariff for the row from the Torquay document. If there is no ad valorem tariff for the row, then leave it blank. As an example, I enter 10% as 10.
----------------

## Schedule Numbering

Schedule | Category | Tariff Act of 1930 Paragraph Number 
-------- | -------- | ----------------------------------
1 | Chemicals, Oils, and Paints | 1 to 97
2 | Earths, Earthenware, and Glassware | 201 to 236
3 | Metals and Manufactures of | 301 to 398
4 | Wood and Manufactures of | 401 to 412
5 | Sugar, Molasses, and Manufactures of | 501 to 506
6 | Tobacco and Manufactures of | 601 to 605
7 | Agricultural Products and Provisions | 701 to 783
8 | Spirits, Wines, and Other Beverages | 801 to 815
9 | Cotton Manufactures | 901 to 924
10 | Flax, Hemp, Jute, and Manufactures of | 1001 to 1022
11 | Wool and Manufactures of | 1101 to 1122
12 | Silk Manufactures | 1201 to 1211
13 | Manufactures of Rayon or Other Synthetic Textile | 1301 to 1313
14 | Papers and Books | 1401 to 1413
15 | Sundries | 1501 to 1559
16 | Title II - Free List | 1601 to 1814
Note: There does not appear to be any Paragraphs in the 100s. I think this is due to the formatting of Schedule 1. In the Tariff Act of 1930 it lists Paragraphs 1 to 97, but I think it is really implying that these are Paragraphs 101 to 197.

### Units Formatting
Below is a table of the numbering used in the Units column and the corresponding meaning. Please add any additional units when necessary.

Number | Meaning
- | -
1 | Cents per pound.
2 | Cents per gallon.
3 | Dollars per pound.
4 | Cents per ounce.
5 | Dollars per ounce.
6 | Cents per square foot.
7 | Cents per ton.
8 | Dollars per ton.
9 | Dollars per gallon.
10 | Cents per cubic foot.
---------------

# Cleaning Data with R

The excel file needs to first be saved as a tab delimited text file ("Geneva, Annecy, Torquay All Schedules.txt"). This R code will take the .txt file and convert it to a clean format. This involves going from a wide to long format, filling in missing tariff values with zeros, and creating variables that will help with future analysis. To run R code make sure the text file is the correct name, the file is in text format, and the working directory is correct. The cleaned file that will be produced is called "Geneva, Annecy, Torquay All Schedules Cleaned.txt".

This same R code is also located in the R Script file called "Clean Tariff Data" in the folder "G:\\MAX-Filer\\Collab\\SOP-kbuzard-S13\\Admin\\torquay". 

Note: This code has been checked with a smaller subset of the entire sample. May want to check that program is still correct with more data.

```{r, eval=FALSE}
library(tidyverse)

#Deleting previous datasets.
rm(list = ls())

# Set working directory. Need to update for your computer.
setwd("~/Desktop/Buzard Update 2020")

#Importing data. Make sure to save "Geneva, Annecy, Torquay All Schedules" excel file as a text file.
data_import <- read_tsv("Geneva, Annecy, Torquay All Schedules.txt")

#Create id.
data_import <- data_import %>%
  mutate(id = row_number())

#Fill in Paragraph numbers using the previous value.
data_import <- data_import %>%
  fill(Paragraph)

#Create id for products within each paragraph. 
data_import <- data_import %>%
  group_by(Paragraph) %>%
  mutate(Product = row_number())

#Creating function for separate datasets for reshaping. Order of the gatt rounds in the select function matters: put from earliest to latest rounds.
separate_table <- function(x_1, x_2, x_3) {
  data_import %>%
    select(id, Schedule, Paragraph, Product, Description, Extra, x_1, x_2, x_3) %>%
    rename(Geneva1 = x_1, Annecy = x_2, Torquay = x_3)
}

#Creating separate datasets for reshaping.
table_specific <- separate_table("Specific_Geneva", "Specific_Annecy", "Specific_Torquay")
table_units <- separate_table("Units_Geneva", "Units_Annecy", "Units_Torquay")
table_ad_valorem <- separate_table("Ad_Valorem_Geneva", "Ad_Valorem_Annecy", "Ad_Valorem_Torquay")

#Reshaping the separate datasets.
tidy_specific <- table_specific %>%
  gather(Geneva1, Annecy, Torquay, key="Gatt_Round", value="Specific")
tidy_units <- table_units %>%
  gather(Geneva1, Annecy, Torquay, key="Gatt_Round", value="Units")
tidy_ad_valorem <- table_ad_valorem %>%
  gather(Geneva1, Annecy, Torquay, key="Gatt_Round", value="Ad_Valorem")

#Combining the separate datasets from above into one dataset.
tidy_gatt <- left_join(tidy_specific, tidy_units) %>%
  left_join(tidy_ad_valorem)

#Create paragraph_product identifier for each observation.
tidy_gatt <- tidy_gatt %>%
  unite(Paragraph_Product, Paragraph, Product, remove = F)

#Fill in observations function. Important that GATT Rounds are in order (i.e. Geneva1, Annecy, Torquay).
fill_missing <- function(x) {
  tidy_gatt %>%
    group_by(Paragraph_Product) %>%
    fill(x)
}

#Fill in observations.
tidy_gatt <- fill_missing("Specific")
tidy_gatt <- fill_missing("Units")
tidy_gatt <- fill_missing("Ad_Valorem")

#Fill in missing with zeros. Can change if you do not want to fill in with zeros.
tidy_gatt <- tidy_gatt %>%
  mutate(Specific = coalesce(Specific, 0)) %>%
  mutate(Units = coalesce(Units, 0)) %>%
  mutate(Ad_Valorem = coalesce(Ad_Valorem, 0))

#Order data. Maybe could use arrange command instead?
tidy_gatt <- tidy_gatt[order(tidy_gatt$Paragraph, tidy_gatt$Product),]

#Removing not needed data.
rm(table_ad_valorem, table_specific, table_units, tidy_ad_valorem, tidy_specific, tidy_units)

#Unique identifier for each observation.
tidy_gatt <- tidy_gatt %>%
  group_by(Paragraph_Product) %>%
  mutate(Round = row_number()) %>%
  unite(Paragraph_Product_Round, Paragraph, Product, Round, remove = F)

#Order the columns.
tidy_gatt <- tidy_gatt %>%
  select(id, Schedule, Paragraph_Product_Round, Paragraph_Product, Paragraph, Product, Round, everything())

View(tidy_gatt)

#Saving cleaned data.
tidy_gatt %>%
  write_tsv("Geneva, Annecy, Torquay All Schedules Cleaned.txt")
```

Below find a table with the variables that will present in the "Geneva, Annecy, Torquay All Schedules Cleaned.txt" file.

Variables | Description
------------ | ---------------
id | Numbering of where the product appears in the excel file.
Schedule | Schedule number from the Tariff Act of 1930.
Paragraph_Product_Round | ID variable that combines Paragraph, Product, and Round variables.
Paragraph_Product | ID variable that combines Paragraph and Product variables.
Paragraph | Tariff Act of 1930 paragraph number.
Product | Numbering of where the product appears within the paragraph.
Round | GATT round number: =1 is Geneva1, =2 is Annecy; =3 Torquay.
Description | Description of products.
Extra | Additional information.
Gatt_Round | Name of GATT round: Geneva1; Annecy; Torquay.
Specific | Specific tariff.
Units | Units on specific tariff.
Ad_Valorem | Ad valorem tariff. Enter 10% as 10.
----------------------

## Tariff Changes with R

The below R code creates variables that identify in which round the tariff changed and in how many rounds it changed. It also identifies which tariffs have both specific and ad valorem components. The above cleaning file needs to be run first to make "Geneva, Annecy, Torquay All Schedules Cleaned.txt". This program takes that file and produces "Geneva, Annecy, Torquay All Schedules Changes.txt".

This same R code is also located in the R Script file called "Tariff Changes" in the folder "G:\\MAX-Filer\\Collab\\SOP-kbuzard-S13\\Admin\\torquay".

Note: This code has been checked with a smaller subset of the entire sample. May want to check that program is still correct with more data.

```{r, eval=FALSE}
library(tidyverse)

#Deleting previous datasets.
rm(list = ls())

# Set working directory. Need to update for your computer.
setwd("~/Desktop/Buzard Update 2020")

#Importing data.
tidy_gatt <- read_tsv("Geneva, Annecy, Torquay All Schedules Cleaned.txt")

#Identify changes in tariffs. Identify combination tariffs (ones with specific and ad valorem)
tidy_gatt <- tidy_gatt %>%
  group_by(Paragraph_Product) %>%
  mutate(Specific_Change = if_else(Specific != lag(Specific), 1, 0, missing = 0)) %>%
  mutate(Specific_Change_Total = sum(Specific_Change)) %>%
  mutate(Ad_Valorem_Change = if_else(Ad_Valorem != lag(Ad_Valorem), 1, 0, missing = 0)) %>%
  mutate(Ad_Valorem_Change_Total = sum(Ad_Valorem_Change)) %>%
  mutate(Tariff_Combination = if_else((Specific > 0) & (Ad_Valorem > 0), 1, 0))

View(tidy_gatt)

#Saving cleaned data.
tidy_gatt %>%
  write_tsv("Geneva, Annecy, Torquay All Schedules Changes.txt")
```

Below find a table with the new variables that will present in the "Geneva, Annecy, Torquay All Schedules Changes.txt" file.

Variables | Description
------------ | ---------------
Specific_Change | =1 if specific tariff changed in this round. =0 otherwise.
Specific_Change_Total | Total number of times the specific tariff changed across all rounds for this product.
Ad_Valorem_Change | =1 if ad valorem tariff changed in this round. =0 otherwise.
Ad_Valorem_Change_Total | Total number of times the ad valorem tariff changed across all rounds for this product.
Tariff_Combination | =1 if product has a specific and ad valorem tariff in the same round. =0 otherwise.
----------------------